### 2022-06-19 analysis/article-results.Rmd
Better version of cal plot

---
title: "Article results"
author: "E. F. Bonneville"
date: "`r Sys.setenv(LANG = 'en_US.UTF-8'); format(Sys.Date(), '%d %B %Y')`"
output: 
  html_document:
    df_print: kable
    toc: yes
    toc_float: 
      smooth_scroll: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(
  echo = FALSE, 
  out.width = "100%", 
  warning = FALSE, 
  message = FALSE
)

# Load libraries for knitting
library(ggplot2)
library(targets)
library(tidyverse)
library(tableone)

# Source functions
source("data-raw/prepare_raw_data.R")
source("R/imputation-helpers.R")
source("R/model-validation-helpers.R")

# Set contrasts for ordered factors
options(contrasts = rep("contr.treatment", 2))

# Set global setting
theme_set(theme_minimal(base_size = 14))
```

```{r objs}
tar_load(
  c(
    dat_combined,
    dat_to_impute,
    imps_all,
    candidate_predictors,
    model_formula,
    analysis_settings,
    validation_dev_lambdamin
  )
)

# For later use
response_var <- all.vars(stats::update(model_formula, . ~ 1))
```

## Descriptives

### Table 1

For continous variables:

```{r table-one-contin}
# Gives you everything for tables 1 and 2 (warnings due to variables not in validation, ignore!)
table_one <- dat_combined %>%  
  select(-all_of(c("sidescored", "AngleCcaIca", "ICAEAtherosclerosis", "StudySubjectID"))) %>%
  # Here we dichotomise purely for the table, does not affect original data
  mutate(
    ICA_nr_90orLarger_geq1 = factor(ifelse(ICA_nr_90orLarger >= 1, 1, 0)),
    ICA_nr_90orLarger_geq2 = factor(ifelse(ICA_nr_90orLarger >= 2, 1, 0)),
    ICAE_NASCET_99 = factor(ifelse(ICAE_NASCET_Degree < 99, 0,1)),
    InnCca_90orLarger_geq1 = factor(ifelse(InnCca_nr_90orLarger >= 1, 1, 0)),
    InnCca_90orLarger_geq2 = factor(ifelse(InnCca_nr_90orLarger >= 2, 1, 0)),
    ICAI_stenosis50 = factor(ifelse(ICAIAtherosclerosis == "Yes, >=50% stenosis", 1, 0)),
    AngleAaInn_OR_AaCca_dich45 = factor(ifelse(AngleAaInn_OR_AaCca > 45, 0, 1))
  ) %>% 
  CreateTableOne(data = ., strata = "dataset")

# Summaries of all continuous variables
kableone(table_one$ContTable, nonnormal = TRUE)
```

For categorical variables:

```{r table-one-categ}
# Summaries of all categorical variables
kableone(table_one$CatTable)
```

### Missing data summaries

```{r miss-dat-summ}
# Seperate datasets
dat_develop <- subset(x = dat_to_impute, select = -dataset, subset = (dataset == "develop"))
dat_valid <- subset(x = dat_to_impute, select = -dataset, subset = (dataset == "valid"))

# Check proportion of complete cases in target analysis (quite good, compare results)
mean(complete.cases(dat_develop[, c("Mc_FailedFemoralApproach", candidate_predictors)]))
mean(complete.cases(dat_valid[, c("Mc_FailedFemoralApproach", candidate_predictors)]))
mean(complete.cases(dat_to_impute[, c("Mc_FailedFemoralApproach", candidate_predictors)]))

# Visualise missings beforehand
naniar::gg_miss_var(dat_to_impute, facet = dataset, show_pct = TRUE)
```

(Naniar/JointAI summaries/VIM)

## Univariable analyses

```{r univar-analyses}
univariate_analyses <- imps_all %>%

  # Only use separately imputed develop, valid and combined cohorts
  filter(imps_label != "imps_assess") %>%

  # Split each of three cohorts, and run analyses on each
  split(f = .$imps_label) %>% 
  map(.f = ~ {
    imps_univariate_ORs(
      outcome = "Mc_FailedFemoralApproach", 
      imps = split(.x, .x$.imp),
      predictors = c(
        candidate_predictors, 
        "ICA_nr_90orLarger",
        "OrigoStenosis50percent",
        "ICAIAtherosclerosis",
        "AngleAaInn_OR_AaCca"
      )
    )
  }) %>% 
  bind_rows(.id = "cohort")

# Make the table
univariate_table <- univariate_analyses %>% 
  select(cohort, term, estimate, `2.5 %`, `97.5 %`) %>%
  mutate(across(where(is.numeric), ~ round(.x, 2))) %>% 
  mutate(
    OR = paste0(estimate, " [", `2.5 %`, "-", `97.5 %`, "]"),
    cohort = factor(
      cohort, 
      levels = c("imps_dev", "imps_valid", "imps_combined"),
      labels = c("develop_cohort", "validation_cohort", "combined_cohort")
    )
  ) %>%

  # Spread and set column order
  pivot_wider(names_from = cohort, values_from = OR, id_cols = term) %>% 
  select(term, develop_cohort, validation_cohort, combined_cohort)

knitr::kable(univariate_table)
```

### Interaction analyses

```{r inters-univar}
# Cohort interactions (Should probably be ommitted from manuscript)
dat_interactions <- imps_all %>% 
  filter(imps_label %in% c("imps_dev", "imps_valid"))

interactions_table <- imps_univariate_ORs(
  outcome = "Mc_FailedFemoralApproach",
  imps = split(x = dat_interactions, f = dat_interactions$.imp),
  predictors = paste0(candidate_predictors, " * imps_label") # imp_label is surrogate for cohort
) %>%

  # Keep only interaction terms - keep  p-values
  filter(grepl(x = term, pattern = "\\:")) %>% 
  select(term, p.value)

knitr::kable(interactions_table)
```

## Internal validation development set

```{r dev-coefs}
# Development imputation
imps_develop <- imps_all %>% 
  filter(imps_label == "imps_assess" & dataset == "develop")

# Development model
mod_develop <- validation_dev_lambdamin$model_fit
mod_coefs <- as.matrix(coef(mod_develop))
colnames(mod_coefs) <- "Coefficients"

knitr::kable(data.frame(mod_coefs))
```

```{r valid-dev}
# Optimism correctedL
validation_dev_lambdamin$validation_summary %>% 
  filter(measure %in% c("auc", "intercept", "slope")) %>% 
  group_by(measure) %>% 
  summarise(value = paste0(corrected, " [", lower_corrected, ";", upper_corrected, "]")) %>%
  knitr::kable()
```

```{r dev-calib}
X_dev <- subset(x = stats::model.matrix(model_formula, data = imps_develop), select = -`(Intercept)`)
y_dev <- imps_develop[[response_var]]
lp_dev <- drop(predict(mod_develop, newx = X_dev, type = "link")) # linear predictor

# Make df for calibration
df_calplot_develop <- cbind.data.frame(lp_dev, imps_develop)

# We use restricted cubic splines with 4 knots
knots <- 4

# Get limits of predicted probabilities across all imputed datasets
pred_probs <- plogis(df_calplot_develop$lp_dev) 
probs_grid <- seq(min(pred_probs), max(pred_probs), by = 0.005)
lp_grid <- qlogis(probs_grid)

# The "average" smooth calibration is simply the one fitted on stacked data
mod_cal_stacked <- glm(
  formula = Mc_FailedFemoralApproach ~ ns(lp_dev, knots), 
  data = df_calplot_develop, 
  family = binomial()
)

# No easy way to create CI for this, so we visualise all the calibration curves
# in each imputed dataset:
split(df_calplot_develop, ~ .imp) %>%
  map(
    .f = ~ {
      mod_cal <- glm(
        formula = Mc_FailedFemoralApproach ~ ns(lp_dev, knots), 
        data = .x, 
        family = binomial()
      )
      data.frame(
        "predicted" = probs_grid,
        "observed" = predict(mod_cal, newdata = data.frame("lp_dev" = lp_grid), type = "response"),
        row.names = NULL
      )
    }
  ) %>%
  bind_rows(.id = ".imp") %>%
  ggplot() +
  geom_line(aes(group = .imp, x = predicted, y = observed), alpha = 0.8, col = "lightgray", size = 1) +
  theme(panel.grid.major = element_blank()) + 
  geom_line(
    data = cbind.data.frame(
      "predicted" = probs_grid,
      "observed" = predict(
        mod_cal_stacked, 
        newdata = data.frame("lp_dev" = lp_grid), 
        type = "response"
      ) #- 0.25
    ),
    aes(x = predicted, y = observed),
    size = 2
  ) +
  geom_abline(intercept = 0, slope = 1, col = "red", size = 1, linetype = "dashed") #+
  geom_histogram(
    data = df_calplot_develop, 
    aes(x = plogis(lp_dev), 
        y = stat(
          0.25 * count / max(count)
        )#0.5 * (..count../sum(..count..))
      ),
    fill = "lightblue",
    bins = 100
  )

# Work this later..

# Add hists of pred probs/annotation of AUC, slope and int?
df_calplot_develop |> 
  ggplot(aes(plogis(lp_dev))) +
  geom_histogram()
```

## External validation

```{r ext-validation}
imps_valid <- imps_all %>% 
  filter(imps_label == "imps_assess" & dataset == "valid")

X_valid <- subset(x = stats::model.matrix(model_formula, data = imps_valid), select = -`(Intercept)`)
y_valid <- imps_valid[[response_var]]

# Compute linear predictor
lp_valid <- drop(predict(mod_develop, newx = X_valid))
df_calplot_valid <- cbind.data.frame(lp_valid, imps_valid)

# Pool AUC
auc_df <- split(df_calplot_valid, ~ .imp) %>%
  map(.f = ~ {
    auc_obj <- pROC::auc(Mc_FailedFemoralApproach ~ lp_valid, data = .x)
    cbind.data.frame("auc" = as.numeric(auc_obj), "var_auc" = pROC::var(auc_obj))
  }) %>%
  bind_rows(.id = ".imp")

psfmi::pool_auc(auc_df$auc, sqrt(auc_df$var_auc), nimp = max(as.numeric(auc_df$.imp)))

# Get cal slope and intercept
split(df_calplot_valid, ~ .imp) |> 
  map(.f = ~ glm(Mc_FailedFemoralApproach ~ lp_valid, family = binomial, data = .x)) |> 
  pool() |> 
  summary()

split(df_calplot_valid, ~ .imp) |> 
  map(.f = ~ glm(Mc_FailedFemoralApproach ~ offset(lp_valid), family = binomial, data = .x)) |> 
  pool() |> 
  summary()

#
pred_probs <- plogis(df_calplot_valid$lp_valid)
probs_grid <- seq(min(pred_probs), max(pred_probs), by = 0.01)
lp_grid <- qlogis(probs_grid)

mod_cal_stacked_valid <- glm(
  formula = Mc_FailedFemoralApproach ~ ns(lp_valid, knots), 
  data = df_calplot_valid, 
  family = binomial()
)

# Now calplot
split(df_calplot_valid, ~ .imp) %>%
  map(
    .f = ~ {
      mod_cal <- glm(
        formula = Mc_FailedFemoralApproach ~ ns(lp_valid, knots), 
        data = .x, 
        family = binomial()
      )
      data.frame(
        "predicted" = probs_grid,
        "observed" = predict(mod_cal, newdata = data.frame("lp_valid" = lp_grid), type = "response"),
        row.names = NULL
      )
    }
  ) %>%
  bind_rows(.id = ".imp") %>%
  ggplot() +
  geom_line(aes(group = .imp, x = predicted, y = observed), alpha = 0.8, col = "lightgray", size = 1) +
  theme(panel.grid.major = element_blank()) + 
  geom_line(
    data = cbind.data.frame(
      "predicted" = probs_grid,
      "observed" = predict(
        mod_cal_stacked_valid, 
        newdata = data.frame("lp_valid" = lp_grid), 
        type = "response"
      )
    ),
    aes(x = predicted, y = observed),
    size = 2
  ) +
  geom_abline(intercept = 0, slope = 1, col = "red", size = 1, linetype = "dashed") +
  lims(x = c(0, 0.8), y = c(0, 0.8))
```

## Updating to combined cohort

```{r combined-validation}

```

### Nomogram

### 2022-06-19 R/calplot-MI.R

@@ -2,12 +2,12 @@ tar_load(
   c(
     imps_all,
     candidate_predictors,
-    validation_dev_min,
+    validation_dev_lambdamin,
     model_formula
   )
 )
 
-model <- validation_dev_min$model_fit
+model <- validation_dev_lambdamin$model_fit
 imps_long <- imps_all %>% 
   filter(imps_label == "imps_assess" & dataset == "develop")
 
@@ -34,6 +34,16 @@ calplot_MI <- function(imps_long,
   pred_probs <- plogis(df_calplot$lp) 
   probs_grid <- seq(min(pred_probs), max(pred_probs), by = 0.005)
   lp_grid <- qlogis(probs_grid)
+  
+  # Plot deciles predicted risks
+  risk_deciles <- quantile(pred_probs, probs = seq(0, 1, by = 0.1))
+  risk_groups <- cut(pred_probs, breaks = risk_deciles, include.lowest = T)
+  
+  # This is wrong!!
+  risk_pts <- cbind.data.frame(
+    "predicted" = tapply(pred_probs, risk_groups, function(x) mean(x, na.rm = TRUE)),
+    "observed" = tapply(as.numeric(y) - 1L, risk_groups, function(x) mean(x, na.rm = TRUE))
+  )
   
   # The "average" smooth calibration is simply the one fitted on stacked data
   smooth_form <- reformulate(response = response_var, termlabels = paste0("splines::ns(lp, ", knots, ")"))
@@ -70,7 +80,8 @@ calplot_MI <- function(imps_long,
       aes(x = predicted, y = observed),
       size = 2
     ) +
-    geom_abline(intercept = 0 - startpos_hist, slope = 1, col = "red", size = 1, linetype = "dashed") +
+    geom_abline(intercept = 0 - startpos_hist,
+                slope = 1, col = "red", size = 1, linetype = "dashed") +
     geom_histogram(
       data = data.frame("predicted" = plogis(df_calplot$lp)),
       aes(x = predicted, y = stat(height_hist * count / max(count))),
@@ -86,14 +97,41 @@ calplot_MI <- function(imps_long,
       ylim = c(ylim[1] - startpos_hist, ylim[2] - startpos_hist),
     )
   
-  return(p)
+  p +
+    geom_point(
+      data = risk_pts,
+      aes(predicted, observed - startpos_hist),
+      size = 2.5,
[…]

### 2022-06-19 R/imputation-helpers.R

@@ -49,6 +49,7 @@ run_imputations <- function(dat_to_impute,
   
   # For ordered covariates to be safe (not sure if Targets passes this on)
   options(contrasts = rep("contr.treatment", 2))
+  type <- match.arg(type)
   
   # Read-in valid and develop separately
   dat_develop <- subset(x = dat_to_impute, select = -dataset, subset = (dataset == "develop"))

### 2022-06-19 _targets.R

@@ -27,7 +27,8 @@ tar_option_set(packages = project_pkgs, error = "continue")
 # Uncomment if running scripts interactively:
 # sapply(project_pkgs, function(pkg) require(pkg, character.only = TRUE)); rm(project_pkgs)
 
-list(plan(callr), plan(callr))# nest for furrr part
+# See https://github.com/ropensci/targets/discussions/359
+plan(list(tweak(callr, workers = 3), tweak(callr, workers = 2)))
 
 
 # Analysis pipeline -------------------------------------------------------
@@ -114,38 +115,27 @@ targets_list <- list(
   ),
   # # Internal validation development set
   tar_target(
-    validation_dev_min,
+    validation_dev_lambdamin,
     validate_lasso_stackedImps(
       imputations = imps_all %>% filter(imps_label == "imps_assess" & dataset == "develop"),
       formula = model_formula,
       wts = "wts",
       n_folds = analysis_settings$n_folds,
-      lambda_choice = "min", #"1se", # crazy results with 1se
-      B = 10 #analysis_settings$B
+      lambda_choice = "min", #"1se" crazy results with 1se, skip
+      B = analysis_settings$B
     )
   ),
   tar_target(
-    validation_dev_1se,
+    validation_comb_lambdamin,
     validate_lasso_stackedImps(
-      imputations = imps_all %>% filter(imps_label == "imps_assess" & dataset == "develop"),
+      imputations = imps_all %>% filter(imps_label == "imps_combined"),
       formula = model_formula,
       wts = "wts",
       n_folds = analysis_settings$n_folds,
-      lambda_choice = "1se", #"1se", # crazy results with 1se
-      B = 10 #analysis_settings$B
+      lambda_choice = "min",
+      B = analysis_settings$B
     )
-  )#,
-  # tar_target(
-  #   fit_validation_combined,
-  #   validate_lasso_stackedImps(
-  #     imputations = imps_all %>% filter(imps_label == "imps_combined"),
-  #     formula = model_formula,
-  #     wts = "wts",
-  #     n_folds = analysis_settings$n_folds,
-  #     lambda_choice = "min",
-  #     B = analysis_settings$B
-  #   )
-  # )
+  )
   # 
   #tarchetypes::tar_render(analysis_summary, path = "analysis/2020-09_analysis-summary.Rmd")
 )

### 2022-06-19 run-pipeline.R

@@ -3,5 +3,6 @@ tar_visnetwork()
 
 # Run/reproduce full analysis
 #tar_make()
+tar_make_future(workers = availableCores())
 
 #tar_meta(fields = warnings)

### 2022-06-19 analysis/article-results.Rmd
Definitively stick with lambda min, 1se too poor

@@
 source("R/model-validation-helpers.R")
+source("R/calplot-MI.R")
 
@@
     analysis_settings,
-    validation_dev_lambdamin
+    validation_dev_lambdamin,
+    validation_comb_lambdamin
   )
@@
 ```{r dev-calib}
-X_dev <- subset(x = stats::model.matrix(model_formula, data = imps_develop), select = -`(Intercept)`)
-y_dev <- imps_develop[[response_var]]
-lp_dev <- drop(predict(mod_develop, newx = X_dev, type = "link")) # linear predictor
-
-# Make df for calibration
-df_calplot_develop <- cbind.data.frame(lp_dev, imps_develop)
-
-# We use restricted cubic splines with 4 knots
-knots <- 4
-
-# Get limits of predicted probabilities across all imputed datasets
-pred_probs <- plogis(df_calplot_develop$lp_dev) 
-probs_grid <- seq(min(pred_probs), max(pred_probs), by = 0.005)
-lp_grid <- qlogis(probs_grid)
-
-# The "average" smooth calibration is simply the one fitted on stacked data
-mod_cal_stacked <- glm(
-  formula = Mc_FailedFemoralApproach ~ ns(lp_dev, knots), 
-  data = df_calplot_develop, 
-  family = binomial()
-)
-
-# No easy way to create CI for this, so we visualise all the calibration curves
-# in each imputed dataset:
-split(df_calplot_develop, ~ .imp) %>%
-  map(
-    .f = ~ {
-      mod_cal <- glm(
-        formula = Mc_FailedFemoralApproach ~ ns(lp_dev, knots), 
-        data = .x, 
-        family = binomial()
-      )
-      data.frame(
-        "predicted" = probs_grid,
-        "observed" = predict(mod_cal, newdata = data.frame("lp_dev" = lp_grid), type = "response"),
-        row.names = NULL
-      )
-    }
-  ) %>%
-  bind_rows(.id = ".imp") %>%
-  ggplot() +
-  geom_line(aes(group = .imp, x = predicted, y = observed), alpha = 0.8, col = "lightgray", size = 1) +
-  theme(panel.grid.major = element_blank()) + 
-  geom_line(
-    data = cbind.data.frame(
-      "predicted" = probs_grid,
-      "observed" = predict(
-        mod_cal_stacked, 
-        newdata = data.frame("lp_dev" = lp_grid), 
-        type = "response"
-      ) #- 0.25
-    ),
-    aes(x = predicted, y = observed),
-    size = 2
+calplot_MI(
+  imps_long = imps_all %>% 
+    filter(imps_label == "imps_assess" & dataset == "develop"),
+  model = validation_dev_lambdamin$model_fit,
+  impdat_ind = ".imp",
+  model_formula = model_formula,
+  height_hist = 0.2,
+  startpos_hist = -0.25,
+  knots = 5,
+  n_bins = 100,
+  xlim = c(0, 0.85),
+  ylim = c(-0.25, 1)
 ) +
-  geom_abline(intercept = 0, slope = 1, col = "red", size = 1, linetype = "dashed") #+
-  geom_histogram(
-    data = df_calplot_develop, 
-    aes(x = plogis(lp_dev), 
-        y = stat(
-          0.25 * count / max(count)
-        )#0.5 * (..count../sum(..count..))
-      ),
-    fill = "lightblue",
-    bins = 100
-  )
-
-# Work this later..
-
-# Add hists of pred probs/annotation of AUC, slope and int?
-df_calplot_develop |> 
-  ggplot(aes(plogis(lp_dev))) +
-  geom_histogram()
+  theme_minimal()
 ```
@@
 X_valid <- subset(x = stats::model.matrix(model_formula, data = imps_valid), select = -`(Intercept)`)
-y_valid <- imps_valid[[response_var]]
 
@@
 
-#
-pred_probs <- plogis(df_calplot_valid$lp_valid)
-probs_grid <- seq(min(pred_probs), max(pred_probs), by = 0.01)
-lp_grid <- qlogis(probs_grid)
+calplot_MI(
+  imps_long = imps_valid,
+  model = validation_dev_lambdamin$model_fit,
+  impdat_ind = ".imp",
+  model_formula = model_formula,
+  height_hist = 0.2,
+  startpos_hist = -0.25,
+  knots = 5,
+  n_bins = 200,
+  xlim = c(0, 0.85),
+  ylim = c(-0.25, 1)
+) +
+  theme_minimal()
+```
 
-mod_cal_stacked_valid <- glm(
-  formula = Mc_FailedFemoralApproach ~ ns(lp_valid, knots), 
-  data = df_calplot_valid, 
-  family = binomial()
+Try intermediate updating with logreg recalibration
+
+```{r updating}
+mod_coefs
+log_recal <- glm(imps_valid$Mc_FailedFemoralApproach ~ lp_valid, family = binomial())
+coef(log_recal)
+
+news_coefs <- c(coef(log_recal)[[1]], coef(log_recal)[[2]] * drop(mod_coefs)[-1])
+new_lp <- drop((news_coefs) %*% t(cbind(1, X_valid)))
+
+df_calplot <- cbind.data.frame("lp" = new_lp, imps_valid)
+knots <- 4
+smooth_form <- reformulate(response = response_var, termlabels = paste0("splines::ns(lp, ", knots, ")"))
+mod_cal_stacked <- glm(formula = smooth_form, data = df_calplot, family = binomial())
+stacked_df <- data.frame(
+    "predicted" = unique(plogis(new_lp)),
+    "observed" = predict(mod_cal_stacked, 
+                         newdata = data.frame("lp" = unique(plogis(new_lp))), 
+                         type = "response")
 )
-
-# Now calplot
-split(df_calplot_valid, ~ .imp) %>%
-  map(
-    .f = ~ {
-      mod_cal <- glm(
-        formula = Mc_FailedFemoralApproach ~ ns(lp_valid, knots), 
-        data = .x, 
-        family = binomial()
-      )
-      data.frame(
-        "predicted" = probs_grid,
-        "observed" = predict(mod_cal, newdata = data.frame("lp_valid" = lp_grid), type = "response"),
-        row.names = NULL
-      )
-    }
-  ) %>%
-  bind_rows(.id = ".imp") %>%
-  ggplot() +
-  geom_line(aes(group = .imp, x = predicted, y = observed), alpha = 0.8, col = "lightgray", size = 1) +
-  theme(panel.grid.major = element_blank()) + 
-  geom_line(
-    data = cbind.data.frame(
-      "predicted" = probs_grid,
-      "observed" = predict(
-        mod_cal_stacked_valid, 
-        newdata = data.frame("lp_valid" = lp_grid), 
-        type = "response"
-      )
-    ),
-    aes(x = predicted, y = observed),
-    size = 2
-  ) +
-  geom_abline(intercept = 0, slope = 1, col = "red", size = 1, linetype = "dashed") +
-  lims(x = c(0, 0.8), y = c(0, 0.8))
+plot(stacked_df$predicted, stacked_df$observed)
 ```
@@
 ```{r combined-validation}
+imps_combined <- imps_all %>% 
+  filter(imps_label == "imps_combined")
 
+mod_combined <- validation_comb_lambdamin$model_fit
+validation_comb_lambdamin$model_fit$lambda
+
+validation_comb_lambdamin$validation_summary
+validation_comb_lambdamin$validation_df |> 
+  ggplot(aes(x = corrected)) +
+  geom_histogram(bins = 50) +
+  facet_wrap(~ measure, scales = "free")
+
+validation_comb_lambdamin$validation_df |> 
+  group_by(boot_num) |> 
+  summarise(pen = mean(penalty)) |>
+  #pull(pen) |>  mean()
+  ggplot(aes(pen)) +
+  geom_histogram()
+
+calplot_MI(
+  imps_long = imps_combined,
+  model = validation_comb_lambdamin$model_fit,
+  impdat_ind = ".imp",
+  model_formula = model_formula,
+  height_hist = 0.2,
+  startpos_hist = -0.25,
+  knots = 5,
+  n_bins = 200,
+  xlim = c(0, 0.5),
+  ylim = c(-0.25, 0.8)
+) +
+  theme_minimal()
+
+# Try the 1se on combined:
+imputations <- imps_combined
+formula <- model_formula
+n_folds = 10
+wts <- "wts"
+lambda_choice <- "1se"
+
+response_var <- all.vars(stats::update(formula, . ~ 1))
+  X_orig <- subset(x = stats::model.matrix(formula, data = imputations), select = -`(Intercept)`)
+  y_orig <- imputations[[response_var]]
+  wts_orig <- imputations[[wts]]
+
+  # First apparent performance
+  folds <- assign_crossval_folds(y = unique(imputations[[".id"]]), n_folds = n_folds)
+  mod_orig <- run_lasso_glmnet(
+    x = X_orig,
+    y = y_orig,
+    wts = wts_orig,
+    foldid = folds[match(imputations[[".id"]], folds[["y"]]), "fold"],
+    lambda_choice = lambda_choice
+  )
+
+mod_orig$lambda
+calplot_MI(
+  imps_long = imps_combined,
+  model = mod_orig,
+  impdat_ind = ".imp",
+  model_formula = model_formula,
+  height_hist = 0.2,
+  startpos_hist = -0.25,
+  knots = 3,
+  n_bins = 200,
+  xlim = c(0, 0.5),
+  ylim = c(-0.25, 0.8)
+) +
+  theme_minimal() 
 ```

### 2022-06-19 R/calplot-MI.R

@@ -1,16 +1,3 @@
-tar_load(
-  c(
-    imps_all,
-    candidate_predictors,
-    validation_dev_lambdamin,
-    model_formula
-  )
-)
-
-model <- validation_dev_lambdamin$model_fit
-imps_long <- imps_all %>% 
-  filter(imps_label == "imps_assess" & dataset == "develop")
-
 calplot_MI <- function(imps_long,
                        impdat_ind = ".imp",
                        model,
@@ -112,33 +99,4 @@ calplot_MI <- function(imps_long,
 }
 
 # See https://stackoverflow.com/questions/35324892/ggplot2-setting-geom-bar-baseline-to-1-instead-of-zero
-calplot_MI(
-  imps_long = imps_all %>% 
-    filter(imps_label == "imps_assess" & dataset == "develop"),
-  model = validation_dev_lambdamin$model_fit,
-  impdat_ind = ".imp",
-  model_formula = model_formula,
-  height_hist = 0.2,
-  startpos_hist = -0.25,
-  knots = 5,
-  n_bins = 100,
-  xlim = c(0, 0.85),
-  ylim = c(-0.25, 1)
-) +
-  theme_minimal()
 
-
-calplot_MI(
-  imps_long = imps_all %>% 
-    filter(imps_label == "imps_assess" & dataset == "valid"),
-  model = validation_dev_lambdamin$model_fit,
-  impdat_ind = ".imp",
-  model_formula = model_formula,
-  height_hist = 0.2,
-  startpos_hist = -0.25,
-  knots = 4,
-  n_bins = 100,
-  xlim = c(0, 0.85),
-  ylim = c(-0.25, 1)
-) +
-  theme_minimal()

### 2022-06-19 _targets.R

@@ -28,7 +28,7 @@ tar_option_set(packages = project_pkgs, error = "continue")
 # sapply(project_pkgs, function(pkg) require(pkg, character.only = TRUE)); rm(project_pkgs)
 
 # See https://github.com/ropensci/targets/discussions/359
-plan(list(tweak(callr, workers = 3), tweak(callr, workers = 2)))
+plan(list(tweak(callr, workers = 1), tweak(callr, workers = 3)))
 
 
 # Analysis pipeline -------------------------------------------------------

### 2022-07-08 analysis/article-results.Rmd
Tweaks to ridge

@@
   map(.f = ~ glm(Mc_FailedFemoralApproach ~ lp_valid, family = binomial, data = .x)) |> 
-  pool() |> 
+  mice::pool() |> 
   summary()
@@
   map(.f = ~ glm(Mc_FailedFemoralApproach ~ offset(lp_valid), family = binomial, data = .x)) |> 
-  pool() |> 
+  mice::pool() |> 
   summary()
@@
 validation_comb_lambdamin$model_fit$lambda
+mod_combined$beta
 
@@
   n_bins = 200,
-  xlim = c(0, 0.5),
+  xlim = c(0, 0.55),
   ylim = c(-0.25, 0.8)
@@
 
-  # First apparent performance
   folds <- assign_crossval_folds(y = unique(imputations[[".id"]]), n_folds = n_folds)
-  mod_orig <- run_lasso_glmnet(
+cross_val <- glmnet::cv.glmnet(
     x = X_orig,
     y = y_orig,
-    wts = wts_orig,
+    weights = wts_orig,
     foldid = folds[match(imputations[[".id"]], folds[["y"]]), "fold"],
-    lambda_choice = lambda_choice
+    family = "binomial", 
+    alpha = 0
   )
 
-mod_orig$lambda
+plot(cross_val)
+
+  # First apparent performance - no penalty..
+mod <-glmnet::glmnet(
+    x = X_orig,
+    y = y_orig,
+    weights = wts_orig,
+    family = "binomial",
+    #lambda = 0, # unpenalized
+    lambda = cross_val$lambda.min,
+    alpha = 0, #ridge
+    intercept = TRUE
+  )
+
+hist(plogis(predict(mod, newx = X_orig)))
+max(plogis(predict(mod, newx = X_orig)))
+coef(mod)
+coef(
+  glm(model_formula, data = imps_combined, family = binomial())
+)
 calplot_MI(
   imps_long = imps_combined,
-  model = mod_orig,
+  model = mod,
   impdat_ind = ".imp",
@@
   startpos_hist = -0.25,
-  knots = 3,
+  knots = 5,
   n_bins = 200,
-  xlim = c(0, 0.5),
-  ylim = c(-0.25, 0.8)
+  xlim = c(0, 1),
+  ylim = c(-0.25, 1)
 ) +
   theme_minimal()
+
+# No penalty
+modo <- glm(
+  model_formula, family = binomial(), data = imps_combined, weights = imps_combined$wts
+)
 ```

### 2022-07-08 R/model-validation-helpers.R

@@ -73,13 +73,14 @@ imps_univariate_ORs <- function(outcome,
 
 
 # Add alpha = 1 to args..
-run_lasso_glmnet <- function(x,
-                             y,
-                             foldid = NULL,
-                             family = "binomial",
-                             lambda_choice = c("min", "1se"),
-                             wts = NULL,
-                             ...) {
+run_stacked_glmnet <- function(x,
+                               y,
+                               foldid = NULL,
+                               family = "binomial",
+                               lambda_choice = c("min", "1se"),
+                               alpha = 0, # ridge
+                               wts = NULL,
+                               ...) {
   
   # Get lambda choice
   lambda_choice <- match.arg(lambda_choice)
@@ -91,6 +92,7 @@ run_lasso_glmnet <- function(x,
     weights = wts,
     foldid = foldid,
     family = family,
+    alpha = alpha,
     ...
   )
   
@@ -104,6 +106,7 @@ run_lasso_glmnet <- function(x,
     weights = wts,
     family = family,
     lambda = lambda,
+    alpha = alpha,
     intercept = TRUE
   )
   
@@ -148,12 +151,13 @@ assess_performance <- function(glmnet_model, # Already with a picked lambda
 }
 
 
-validate_lasso_stackedImps <- function(imputations, # as returned by "long"
-                                       formula,
-                                       wts, # var_name with weights
-                                       n_folds = 10,
-                                       lambda_choice = c("min", "1se"),
-                                       B = 5) {
+validate_stackedImps <- function(imputations, # as returned by "long"
+                                 formula,
+                                 wts, # var_name with weights
+                                 n_folds = 10,
+                                 lambda_choice = c("min", "1se"),
+                                 alpha = 0, # ridge
+                                 B = 5) {
   
   # For ordered covariates to be safe (not sure if Targets passes this on)
   options(contrasts = rep("contr.treatment", 2))
@@ -169,12 +173,13 @@ validate_lasso_stackedImps <- function(imputations, # as returned by "long"
   
   # First apparent performance
   folds <- assign_crossval_folds(y = unique(imputations[[".id"]]), n_folds = n_folds)
-  mod_orig <- run_lasso_glmnet(
+  mod_orig <- run_stacked_glmnet(
     x = X_orig,
[…]

### 2022-07-08 _targets.R

@@ -28,8 +28,8 @@ tar_option_set(packages = project_pkgs, error = "continue")
 # sapply(project_pkgs, function(pkg) require(pkg, character.only = TRUE)); rm(project_pkgs)
 
 # See https://github.com/ropensci/targets/discussions/359
-plan(list(tweak(callr, workers = 1), tweak(callr, workers = 3)))
-
+#plan(list(tweak(callr, workers = 1), tweak(callr, workers = 3)))
+plan(callr)
 
 # Analysis pipeline -------------------------------------------------------
 
@@ -67,11 +67,9 @@ targets_list <- list(
   # -- Part 2: imputations
   
   # We pick the number of imputations and specify our candidate predictors
-  # such that we can directly add the weights for the stacked LASSO
-  tar_target(
-    analysis_settings, 
-    list("m" = 50L, "n_cycles" = 15L, "B" = 200L, "n_folds" = 10L)
-  ),
+  # such that we can directly add the weights for the stacked ridge
+  tar_target(validation_settings, list("B" = 3L, "n_folds" = 10L)), # change back to B = 200L
+  tar_target(imp_settings, list("m" = 50L, "n_cycles" = 15L)),
   tar_target(
     candidate_predictors, c(
       "M_age",
@@ -85,14 +83,14 @@ targets_list <- list(
   ),
   tar_target(
     dat_to_impute, 
-    add_stacked_weights(dat_combined, candidate_predictors, analysis_settings)
+    add_stacked_weights(dat_combined, candidate_predictors, imp_settings)
   ),
   
   # All rounds of imputations here:
-  tar_target(imps_dev, run_imputations(dat_to_impute, analysis_settings, type = "develop")),
-  tar_target(imps_valid, run_imputations(dat_to_impute, analysis_settings, type = "valid")),
-  tar_target(imps_assess, run_imputations(dat_to_impute, analysis_settings, type = "model")),
-  tar_target(imps_combined, run_imputations(dat_to_impute, analysis_settings, type = "combined")),
+  tar_target(imps_dev, run_imputations(dat_to_impute, imp_settings, type = "develop")),
+  tar_target(imps_valid, run_imputations(dat_to_impute, imp_settings, type = "valid")),
+  tar_target(imps_assess, run_imputations(dat_to_impute, imp_settings, type = "model")),
+  tar_target(imps_combined, run_imputations(dat_to_impute, imp_settings, type = "combined")),
   
   # Combine all imputed datasets in one big df
   tar_target(
[…]

### 2022-07-08 run-pipeline.R

@@ -3,6 +3,6 @@ tar_visnetwork()
 
 # Run/reproduce full analysis
 #tar_make()
-tar_make_future(workers = availableCores())
+tar_make_future(workers = availableCores() - 1) # use min(1, ...) formulation
 
 #tar_meta(fields = warnings)

### 2022-07-10 analysis/article-results.Rmd
Tests of simple recalibration

@@
 ```{r updating}
-mod_coefs
-log_recal <- glm(imps_valid$Mc_FailedFemoralApproach ~ lp_valid, family = binomial())
-coef(log_recal)
+# 
+mod_recal <- split(df_calplot_valid, ~ .imp) |> 
+  map(.f = ~ glm(Mc_FailedFemoralApproach ~ lp_valid, family = binomial, data = .x)) |> 
+  mice::pool() |> 
+  summary()
+coefs_mod_recal <- mod_recal$estimate
+new_coefs <- drop(coefs_mod_recal[2] * mod_coefs[-1])
+new_coefs <- c(mod_coefs[1] + coefs_mod_recal[1], new_coefs) #new_coefs[1] - 0.1059395 #+ coefs_mod_recal[1]
+new_lp <- drop((new_coefs) %*% t(cbind(1, X_valid)))
 
-news_coefs <- c(coef(log_recal)[[1]], coef(log_recal)[[2]] * drop(mod_coefs)[-1])
-new_lp <- drop((news_coefs) %*% t(cbind(1, X_valid)))
+library(Hmisc)
+library(rms)
 
-df_calplot <- cbind.data.frame("lp" = new_lp, imps_valid)
+dd <- datadist(imps_valid %>% select(all_of(c(response_var, candidate_predictors)))); 
+options(datadist = "dd")
+
+CalibrationCurves::val.prob.ci.2(
+  y = (as.numeric(imps_valid$Mc_FailedFemoralApproach) - 1)[1:30000],
+  logit = (lp_valid)[1:30000], 
+  smooth = FALSE,
+  CL.smooth	= FALSE,
+  CL.BT	= FALSE,
+  logistic.cal = TRUE,
+  #g = 10,
+  dostats = FALSE
+)
+
+df_calplot <- cbind.data.frame("lp_recal" = new_lp, imps_valid)
 knots <- 4
-smooth_form <- reformulate(response = response_var, termlabels = paste0("splines::ns(lp, ", knots, ")"))
-mod_cal_stacked <- glm(formula = smooth_form, data = df_calplot, family = binomial())
+#smooth_form <- reformulate(response = response_var, termlabels = paste0("splines::ns(lp, ", knots, ")"))
+mod_cal_stacked <- glm(Mc_FailedFemoralApproach ~ ns(lp_recal, 4), data = df_calplot, family = binomial())
 stacked_df <- data.frame(
-    "predicted" = unique(plogis(new_lp)),
+    "predicted" = seq(0.01, 1, by = 0.01),
     "observed" = predict(mod_cal_stacked, 
-                         newdata = data.frame("lp" = unique(plogis(new_lp))), 
+                         newdata = data.frame("lp_recal" = seq(0.01, 1, by = 0.01)), 
                          type = "response")
 )
-plot(stacked_df$predicted, stacked_df$observed)
+plot(stacked_df$predicted, stacked_df$observed, xlim = c(0, 1), ylim = c(0, 1))
+
+# Try other
+
+# Try first all in one imp dataset
+imp1 <- subset(df_calplot_valid, .imp == 1)
+
+# See https://darrendahly.github.io/post/homr/
+m1 <- glm(Mc_FailedFemoralApproach ~ offset(lp_valid), data = imp1, family = binomial())
+m2 <- glm(Mc_FailedFemoralApproach ~ lp_valid, data = imp1, family = binomial())
+imp1$pred_m1 <- predict(m1, type = "response")
+imp1$pred_m2 <- predict(m2, type = "response")
+
+ggplot(data = imp1) +
+  geom_smooth(
+    aes(x = pred_m1, y = as.numeric(Mc_FailedFemoralApproach) - 1), 
+                  color = "red", se = FALSE, method = "loess"
+  ) +
+  geom_smooth(
+    aes(x = plogis(lp_valid), y = as.numeric(Mc_FailedFemoralApproach) - 1), 
+                  color = "black", se = FALSE, method = "loess"
+  ) + 
+  geom_smooth(
+    aes(x = pred_m2, y = as.numeric(Mc_FailedFemoralApproach) - 1), 
+                  color = "blue", se = FALSE, method = "loess"
+  ) + 
+  geom_abline(intercept = 0, slope = 1) +
+  lims(x = c(0, 1), y = c(0, 1))
+
+# Other tests
+
 ```

### 2022-07-10 _targets.R

@@ -68,8 +68,8 @@ targets_list <- list(
   
   # We pick the number of imputations and specify our candidate predictors
   # such that we can directly add the weights for the stacked ridge
-  tar_target(validation_settings, list("B" = 3L, "n_folds" = 10L)), # change back to B = 200L
-  tar_target(imp_settings, list("m" = 50L, "n_cycles" = 15L)),
+  tar_target(validation_settings, list("B" = 200L, "n_folds" = 10L)), 
+  tar_target(imp_settings, list("m" = 50L, "n_cycles" = 20L)),
   tar_target(
     candidate_predictors, c(
       "M_age",

### 2022-07-11 analysis/article-results.Rmd
Unpenalized multivar interaction-by-cohort is better

@@
     validation_dev_lambdamin,
-    validation_comb_lambdamin
+    validation_comb_lambdamin,
+    validation_dev_lambda1se,
+    validation_comb_lambda1se
   )
@@
 naniar::gg_miss_var(dat_to_impute, facet = dataset, show_pct = TRUE)
+naniar::gg_miss_upset(dat_valid, nsets = 10)
+naniar::gg_miss_upset(dat_develop, nsets = 10)
 ```
@@
 ```{r valid-dev}
-# Optimism correctedL
+# Optimism corrected
 validation_dev_lambdamin$validation_summary %>% 
@@
     filter(imps_label == "imps_assess" & dataset == "develop"),
-  model = validation_dev_lambdamin$model_fit,
+  model = mod_develop_1se, #validation_dev_lambdamin$model_fit,
   impdat_ind = ".imp",
@@
 
+# 1 se test
+folds <- assign_crossval_folds(y = unique(imps_develop[[".id"]]), n_folds = 10)
+cross_val <- glmnet::cv.glmnet(
+    x = subset(x = stats::model.matrix(model_formula, data = imps_develop), select = -`(Intercept)`),
+    y = imps_develop$Mc_FailedFemoralApproach,
+    wts = imps_develop$wts,
+    foldid = folds[match(imps_develop[[".id"]], folds[["y"]]), "fold"],
+    family = "binomial",
+    alpha = 0
+  )
+plot(cross_val)
+
+folds <- assign_crossval_folds(y = unique(imps_all[[".id"]]), n_folds = 10)
+
+cross_val <- glmnet::cv.glmnet(
+    x = subset(x = stats::model.matrix(model_formula, data = imps_all), select = -`(Intercept)`),
+    y = imps_all$Mc_FailedFemoralApproach,
+    wts = imps_all$wts,
+    foldid = folds[match(imps_all[[".id"]], folds[["y"]]), "fold"],
+    family = "binomial",
+    alpha = 0
+  )
+cross_val$lambda
+plot(cross_val)
+
+mod_develop_1se <- run_stacked_glmnet(
+    x = subset(x = stats::model.matrix(model_formula, data = imps_develop), select = -`(Intercept)`),
+    y = imps_develop$Mc_FailedFemoralApproach,
+    wts = imps_develop$wts,
+    foldid = folds[match(imps_develop[[".id"]], folds[["y"]]), "fold"],
+    lambda_choice = "1se",
+    alpha = 0
+  )
+
 # Compute linear predictor
-lp_valid <- drop(predict(mod_develop, newx = X_valid))
+lp_valid <- drop(predict(mod_develop_1se, newx = X_valid))
 df_calplot_valid <- cbind.data.frame(lp_valid, imps_valid)
@@
   imps_long = imps_valid,
-  model = validation_dev_lambdamin$model_fit,
+  model = mod_develop_1se, #validation_dev_lambdamin$model_fit,
   impdat_ind = ".imp",
@@
 
+calplot_MI(
+  imps_long = imps_combined,
+  model = validation_comb_lambda1se$model_fit,
+  impdat_ind = ".imp",
+  model_formula = model_formula,
+  height_hist = 0.2,
+  startpos_hist = -0.25,
+  knots = 3,
+  n_bins = 200,
+  xlim = c(0, 0.55),
+  ylim = c(-0.25, 0.8)
+) +
+  theme_minimal()
+
 # Try the 1se on combined:
@@
 
+```{r interaction-by-cohort}
+split(imps_combined, ~ .imp) |> 
+  map(.f = ~ glm(update(model_formula, . ~ . * dataset), family = binomial, data = .x)) |> 
+  mice::pool() |> 
+  broom::tidy() |>  
+  ggforestplot::forestplot(
+    name = term,
+    estimate = estimate,
+    se = std.error
+  )
+
+split(imps_combined, ~ .imp) |> 
+  map(.f = ~ glm(update(model_formula, . ~ . * dataset), family = binomial, data = .x)) |> 
+  mice::pool() |> 
+  broom::tidy() |> 
+  select(term, estimate)
+
+split(imps_combined |> filter(dataset == "develop"), ~ .imp) |> 
+  map(.f = ~ glm(model_formula, family = binomial, data = .x)) |> 
+  mice::pool() |> 
+  broom::tidy() |> 
+  pull(estimate)
+
+cbind.data.frame(
+  split(imps_combined, ~ .imp) |> 
+  map(.f = ~ glm(update(model_formula, . ~ . * dataset), family = binomial, data = .x)) |> 
+  mice::pool() |> 
+  broom::tidy() |> 
+  select(term, estimate),
+  "develop" = split(imps_combined |> filter(dataset == "develop"), ~ .imp) |> 
+  map(.f = ~ glm(model_formula, family = binomial, data = .x)) |> 
+  mice::pool() |> 
+  broom::tidy() |> 
+  pull(estimate),
+  "valid" =  split(imps_combined |> filter(dataset == "valid"), ~ .imp) |> 
+  map(.f = ~ glm(model_formula, family = binomial, data = .x)) |> 
+  mice::pool() |> 
+  broom::tidy() |> 
+  pull(estimate),
+  "combined" =  split(imps_combined, ~ .imp) |> 
+  map(.f = ~ glm(model_formula, family = binomial, data = .x)) |> 
+  mice::pool() |> 
+  broom::tidy() |> 
+  pull(estimate)
+) |> 
+  View()
+
+# ORs
+cbind.data.frame(
+  split(imps_combined, ~ .imp) |> 
+  map(.f = ~ glm(update(model_formula, . ~ . * dataset), family = binomial, data = .x)) |> 
+  mice::pool() |> 
+  broom::tidy(exponentiate = TRUE) |> 
+  select(term, estimate),
+  "develop" = split(imps_combined |> filter(dataset == "develop"), ~ .imp) |> 
+  map(.f = ~ glm(model_formula, family = binomial, data = .x)) |> 
+  mice::pool() |> 
+  broom::tidy(exponentiate = TRUE) |> 
+  pull(estimate),
+  "valid" =  split(imps_combined |> filter(dataset == "valid"), ~ .imp) |> 
+  map(.f = ~ glm(model_formula, family = binomial, data = .x)) |> 
+  mice::pool() |> 
+  broom::tidy(exponentiate = TRUE) |> 
+  pull(estimate),
+  "combined" =  split(imps_combined, ~ .imp) |> 
+  map(.f = ~ glm(model_formula, family = binomial, data = .x)) |> 
+  mice::pool() |> 
+  broom::tidy(exponentiate = TRUE) |> 
+  pull(estimate)
+) |> 
+  View()
+
+model.matrix(model_formula, data=imps_combined) %>% 
+  cor(use="pairwise.complete.obs") %>% 
+  ggcorrplot::ggcorrplot(show.diag = F, type="lower", lab=TRUE, lab_size=2)
+```
+
 ### Nomogram

### 2022-07-11 R/calplot-MI.R

@@ -26,15 +26,28 @@ calplot_MI <- function(imps_long,
   risk_deciles <- quantile(pred_probs, probs = seq(0, 1, by = 0.1))
   risk_groups <- cut(pred_probs, breaks = risk_deciles, include.lowest = T)
   
+  # Scale the wts
+  scaled_wts <- imps_long$wts / sum(imps_long$wts)
+  
+  # https://stats.stackexchange.com/questions/220156/how-to-combine-pool-binomial-confidence-intervals-after-multiple-imputation 
+  #https://github.com/mwheymans/miceafter/blob/main/R/pool_prop_wilson.R
+  
+  # (Also see https://en.wikipedia.org/wiki/Binomial_proportion_confidence_interval#Standard_error_of_a_proportion_estimation_when_using_weighted_data)
   # This is wrong!!
   risk_pts <- cbind.data.frame(
     "predicted" = tapply(pred_probs, risk_groups, function(x) mean(x, na.rm = TRUE)),
-    "observed" = tapply(as.numeric(y) - 1L, risk_groups, function(x) mean(x, na.rm = TRUE))
+    "observed" = tapply(as.numeric(y) - 1L, risk_groups, function(x) mean(x, na.rm = TRUE)),
+    "sum_wts" = tapply(imps_long$wts, risk_groups, function(x) sum(x^2)), #sum((x / sum(x))^2)),
+    "n" = as.numeric(table(risk_groups))
   )
+  risk_pts$se <- with(risk_pts, sqrt(observed * (1 - observed) / n))
+  risk_pts$lower <- with(risk_pts, observed - qnorm(0.975) * se)
+  risk_pts$upper <- with(risk_pts, observed + qnorm(0.975) * se)
   
   # The "average" smooth calibration is simply the one fitted on stacked data
   smooth_form <- reformulate(response = response_var, termlabels = paste0("splines::ns(lp, ", knots, ")"))
-  mod_cal_stacked <- glm(formula = smooth_form, data = df_calplot, family = binomial())
+  mod_cal_stacked <- glm(formula = smooth_form, data = df_calplot, family = binomial(),#)#,
+                         weights = df_calplot$wts)
   stacked_df <- data.frame(
     "predicted" = probs_grid,
     "observed" = predict(mod_cal_stacked, newdata = data.frame("lp" = lp_grid), type = "response") - startpos_hist
@@ -67,8 +80,10 @@ calplot_MI <- function(imps_long,
       aes(x = predicted, y = observed),
[…]

### 2022-07-11 _targets.R

@@ -133,6 +133,30 @@ targets_list <- list(
       lambda_choice = "min",
       B = validation_settings$B
     )
+  ),
+  
+  # Build them with 1se rule
+  tar_target(
+    validation_dev_lambda1se,
+    validate_stackedImps(
+      imputations = imps_all %>% filter(imps_label == "imps_assess" & dataset == "develop"),
+      formula = model_formula,
+      wts = "wts",
+      n_folds = validation_settings$n_folds,
+      lambda_choice = "1se", #"1se" crazy results with 1se, skip
+      B = 3 #validation_settings$B
+    )
+  ),
+  tar_target(
+    validation_comb_lambda1se,
+    validate_stackedImps(
+      imputations = imps_all %>% filter(imps_label == "imps_combined"),
+      formula = model_formula,
+      wts = "wts",
+      n_folds = validation_settings$n_folds,
+      lambda_choice = "1se",
+      B = 3 #validation_settings$B
+    )
   )
   # 
   #tarchetypes::tar_render(analysis_summary, path = "analysis/2020-09_analysis-summary.Rmd")

### 2022-07-11 analysis/article-results.Rmd
Checks with ggally
Remove univariable analyses + remove 1se attempts, start of trimming rmd

@@
     validation_dev_lambdamin,
-    validation_comb_lambdamin,
-    validation_dev_lambda1se,
-    validation_comb_lambda1se
+    validation_comb_lambdamin
   )
@@
 (Naniar/JointAI summaries/VIM)
-
-## Univariable analyses
-
-```{r univar-analyses}
-univariate_analyses <- imps_all %>%
-
-  # Only use separately imputed develop, valid and combined cohorts
-  filter(imps_label != "imps_assess") %>%
-
-  # Split each of three cohorts, and run analyses on each
-  split(f = .$imps_label) %>% 
-  map(.f = ~ {
-    imps_univariate_ORs(
-      outcome = "Mc_FailedFemoralApproach", 
-      imps = split(.x, .x$.imp),
-      predictors = c(
-        candidate_predictors, 
-        "ICA_nr_90orLarger",
-        "OrigoStenosis50percent",
-        "ICAIAtherosclerosis",
-        "AngleAaInn_OR_AaCca"
-      )
-    )
-  }) %>% 
-  bind_rows(.id = "cohort")
-
-# Make the table
-univariate_table <- univariate_analyses %>% 
-  select(cohort, term, estimate, `2.5 %`, `97.5 %`) %>%
-  mutate(across(where(is.numeric), ~ round(.x, 2))) %>% 
-  mutate(
-    OR = paste0(estimate, " [", `2.5 %`, "-", `97.5 %`, "]"),
-    cohort = factor(
-      cohort, 
-      levels = c("imps_dev", "imps_valid", "imps_combined"),
-      labels = c("develop_cohort", "validation_cohort", "combined_cohort")
-    )
-  ) %>%
-
-  # Spread and set column order
-  pivot_wider(names_from = cohort, values_from = OR, id_cols = term) %>% 
-  select(term, develop_cohort, validation_cohort, combined_cohort)
-
-knitr::kable(univariate_table)
-```
-
-### Interaction analyses
-
-```{r inters-univar}
-# Cohort interactions (Should probably be ommitted from manuscript)
-dat_interactions <- imps_all %>% 
-  filter(imps_label %in% c("imps_dev", "imps_valid"))
-
-interactions_table <- imps_univariate_ORs(
-  outcome = "Mc_FailedFemoralApproach",
-  imps = split(x = dat_interactions, f = dat_interactions$.imp),
-  predictors = paste0(candidate_predictors, " * imps_label") # imp_label is surrogate for cohort
-) %>%
-
-  # Keep only interaction terms - keep  p-values
-  filter(grepl(x = term, pattern = "\\:")) %>% 
-  select(term, p.value)
-
-knitr::kable(interactions_table)
-```
 
@@
 
+dat_combined |> 
+  select(candidate_predictors, dataset) |>  
+    #Mc_FailedFemoralApproach, dataset) |> 
+  GGally::ggpairs(
+    aes(col = dataset, fill = dataset)
+  )
+
   # First apparent performance - no penalty..

### 2022-07-11 _targets.R

@@ -133,30 +133,6 @@ targets_list <- list(
       lambda_choice = "min",
       B = validation_settings$B
     )
-  ),
-  
-  # Build them with 1se rule
-  tar_target(
-    validation_dev_lambda1se,
-    validate_stackedImps(
-      imputations = imps_all %>% filter(imps_label == "imps_assess" & dataset == "develop"),
-      formula = model_formula,
-      wts = "wts",
-      n_folds = validation_settings$n_folds,
-      lambda_choice = "1se", #"1se" crazy results with 1se, skip
-      B = 3 #validation_settings$B
-    )
-  ),
-  tar_target(
-    validation_comb_lambda1se,
-    validate_stackedImps(
-      imputations = imps_all %>% filter(imps_label == "imps_combined"),
-      formula = model_formula,
-      wts = "wts",
-      n_folds = validation_settings$n_folds,
-      lambda_choice = "1se",
-      B = 3 #validation_settings$B
-    )
   )
   # 
   #tarchetypes::tar_render(analysis_summary, path = "analysis/2020-09_analysis-summary.Rmd")

### 2022-07-11 analysis/article-results.Rmd
Crappy attempts with geom_smooth

@@
 
-(Naniar/JointAI summaries/VIM)
+<!-- (Naniar/JointAI summaries/VIM) -->
 
@@
 
+# Print penalied model coefficients
 knitr::kable(data.frame(mod_coefs))
 ```
+
+Bootstrap-based internal validation, optimism-corrected performance:
 
@@
 
+Calibration plot (apparent)
+
 ```{r dev-calib}
@@
     filter(imps_label == "imps_assess" & dataset == "develop"),
-  model = mod_develop_1se, #validation_dev_lambdamin$model_fit,
+  model = validation_dev_lambdamin$model_fit,
   impdat_ind = ".imp",
@@
 
-# 1 se test
-folds <- assign_crossval_folds(y = unique(imps_develop[[".id"]]), n_folds = 10)
-cross_val <- glmnet::cv.glmnet(
-    x = subset(x = stats::model.matrix(model_formula, data = imps_develop), select = -`(Intercept)`),
-    y = imps_develop$Mc_FailedFemoralApproach,
-    wts = imps_develop$wts,
-    foldid = folds[match(imps_develop[[".id"]], folds[["y"]]), "fold"],
-    family = "binomial",
-    alpha = 0
-  )
-plot(cross_val)
-
-folds <- assign_crossval_folds(y = unique(imps_all[[".id"]]), n_folds = 10)
-
-cross_val <- glmnet::cv.glmnet(
-    x = subset(x = stats::model.matrix(model_formula, data = imps_all), select = -`(Intercept)`),
-    y = imps_all$Mc_FailedFemoralApproach,
-    wts = imps_all$wts,
-    foldid = folds[match(imps_all[[".id"]], folds[["y"]]), "fold"],
-    family = "binomial",
-    alpha = 0
-  )
-cross_val$lambda
-plot(cross_val)
-
-mod_develop_1se <- run_stacked_glmnet(
-    x = subset(x = stats::model.matrix(model_formula, data = imps_develop), select = -`(Intercept)`),
-    y = imps_develop$Mc_FailedFemoralApproach,
-    wts = imps_develop$wts,
-    foldid = folds[match(imps_develop[[".id"]], folds[["y"]]), "fold"],
-    lambda_choice = "1se",
-    alpha = 0
-  )
-
 # Compute linear predictor
-lp_valid <- drop(predict(mod_develop_1se, newx = X_valid))
+lp_valid <- drop(predict(mod_develop, newx = X_valid))
 df_calplot_valid <- cbind.data.frame(lp_valid, imps_valid)
@@
 
+# Overestimation of risks
 calplot_MI(
   imps_long = imps_valid,
-  model = mod_develop_1se, #validation_dev_lambdamin$model_fit,
+  model = validation_dev_lambdamin$model_fit,
   impdat_ind = ".imp",
@@
 
-Try intermediate updating with logreg recalibration
+## Model update 1: logistic recalibration
 
-```{r updating}
+```{r log-recal}
 # 
-mod_recal <- split(df_calplot_valid, ~ .imp) |> 
-  map(.f = ~ glm(Mc_FailedFemoralApproach ~ lp_valid, family = binomial, data = .x)) |> 
-  mice::pool() |> 
-  summary()
-coefs_mod_recal <- mod_recal$estimate
-new_coefs <- drop(coefs_mod_recal[2] * mod_coefs[-1])
-new_coefs <- c(mod_coefs[1] + coefs_mod_recal[1], new_coefs) #new_coefs[1] - 0.1059395 #+ coefs_mod_recal[1]
-new_lp <- drop((new_coefs) %*% t(cbind(1, X_valid)))
+imps_valid$lp_valid <- lp_valid
+mod_recal <- glm(
+  Mc_FailedFemoralApproach ~ lp_valid, 
+  family = binomial(), 
+  data = imps_valid
+)
+boots_valid <- group_resample(imps_valid, id_var = ".id", 
+                              B = 50, 
+    boot_id_name = ".id_boot")
 
-library(Hmisc)
-library(rms)
+boots_cals <- map_dfr(boots_valid, function(boot_dat) {
+  mod_recal_boot <- glm(
+  Mc_FailedFemoralApproach ~ lp_valid, 
+  family = binomial(), 
+  data = boot_dat
+)
+  calibration_intercept_slope(
+    imps_valid$Mc_FailedFemoralApproach, predict(mod_recal_boot, newdata = imps_valid)
+  )
+})
+colMeans(boots_cals)
 
-dd <- datadist(imps_valid %>% select(all_of(c(response_var, candidate_predictors)))); 
-options(datadist = "dd")
-
-CalibrationCurves::val.prob.ci.2(
-  y = (as.numeric(imps_valid$Mc_FailedFemoralApproach) - 1)[1:30000],
-  logit = (lp_valid)[1:30000], 
-  smooth = FALSE,
-  CL.smooth	= FALSE,
-  CL.BT	= FALSE,
-  logistic.cal = TRUE,
-  #g = 10,
-  dostats = FALSE
-)
-
-df_calplot <- cbind.data.frame("lp_recal" = new_lp, imps_valid)
-knots <- 4
-#smooth_form <- reformulate(response = response_var, termlabels = paste0("splines::ns(lp, ", knots, ")"))
-mod_cal_stacked <- glm(Mc_FailedFemoralApproach ~ ns(lp_recal, 4), data = df_calplot, family = binomial())
-stacked_df <- data.frame(
-    "predicted" = seq(0.01, 1, by = 0.01),
-    "observed" = predict(mod_cal_stacked, 
-                         newdata = data.frame("lp_recal" = seq(0.01, 1, by = 0.01)), 
-                         type = "response")
-)
-plot(stacked_df$predicted, stacked_df$observed, xlim = c(0, 1), ylim = c(0, 1))
-
-# Try other
-
-# Try first all in one imp dataset
-imp1 <- subset(df_calplot_valid, .imp == 1)
-
-# See https://darrendahly.github.io/post/homr/
-m1 <- glm(Mc_FailedFemoralApproach ~ offset(lp_valid), data = imp1, family = binomial())
-m2 <- glm(Mc_FailedFemoralApproach ~ lp_valid, data = imp1, family = binomial())
-imp1$pred_m1 <- predict(m1, type = "response")
-imp1$pred_m2 <- predict(m2, type = "response")
-
-ggplot(data = imp1) +
+# Calib plot
+library(splines)
+ggplot() +
   geom_smooth(
-    aes(x = pred_m1, y = as.numeric(Mc_FailedFemoralApproach) - 1), 
-                  color = "red", se = FALSE, method = "loess"
-  ) +
-  geom_smooth(
-    aes(x = plogis(lp_valid), y = as.numeric(Mc_FailedFemoralApproach) - 1), 
-                  color = "black", se = FALSE, method = "loess"
-  ) + 
-  geom_smooth(
-    aes(x = pred_m2, y = as.numeric(Mc_FailedFemoralApproach) - 1), 
-                  color = "blue", se = FALSE, method = "loess"
+    data = cbind.data.frame(
+      "predicted" = predict(mod_recal, type = "response"), #wrrooong
+      "outcome" = as.numeric(imps_valid$Mc_FailedFemoralApproach) - 1L,
+      "imp" = imps_valid$.imp
+    ),
+    aes(
+      x = predicted,
+      y = outcome,
+      group = imp
+    ), method = "glm",
+    formula = y ~ ns(x, 5),
+    method.args = list(family = "binomial"),
+    se = FALSE,
+    col = "gray",
+    alpha = 0.7
   ) +
@@
   lims(x = c(0, 1), y = c(0, 1))
-
-# Other tests
-
 ```
 
-## Updating to combined cohort
+## Model update 2: combined cohort re-estimation
 
-```{r combined-validation}
-imps_combined <- imps_all %>% 
-  filter(imps_label == "imps_combined")
+### Interaction-by-cohort (unpenalized)
 
-mod_combined <- validation_comb_lambdamin$model_fit
-validation_comb_lambdamin$model_fit$lambda
-mod_combined$beta
-
-validation_comb_lambdamin$validation_summary
-validation_comb_lambdamin$validation_df |> 
-  ggplot(aes(x = corrected)) +
-  geom_histogram(bins = 50) +
-  facet_wrap(~ measure, scales = "free")
-
-validation_comb_lambdamin$validation_df |> 
-  group_by(boot_num) |> 
-  summarise(pen = mean(penalty)) |>
-  #pull(pen) |>  mean()
-  ggplot(aes(pen)) +
-  geom_histogram()
-
-calplot_MI(
-  imps_long = imps_combined,
-  model = validation_comb_lambdamin$model_fit,
-  impdat_ind = ".imp",
-  model_formula = model_formula,
-  height_hist = 0.2,
-  startpos_hist = -0.25,
-  knots = 5,
-  n_bins = 200,
-  xlim = c(0, 0.55),
-  ylim = c(-0.25, 0.8)
-) +
-  theme_minimal()
-
-calplot_MI(
-  imps_long = imps_combined,
-  model = validation_comb_lambda1se$model_fit,
-  impdat_ind = ".imp",
-  model_formula = model_formula,
-  height_hist = 0.2,
-  startpos_hist = -0.25,
-  knots = 3,
-  n_bins = 200,
-  xlim = c(0, 0.55),
-  ylim = c(-0.25, 0.8)
-) +
-  theme_minimal()
-
-# Try the 1se on combined:
-imputations <- imps_combined
-formula <- model_formula
-n_folds = 10
-wts <- "wts"
-lambda_choice <- "1se"
-
-response_var <- all.vars(stats::update(formula, . ~ 1))
-  X_orig <- subset(x = stats::model.matrix(formula, data = imputations), select = -`(Intercept)`)
-  y_orig <- imputations[[response_var]]
-  wts_orig <- imputations[[wts]]
-
-  folds <- assign_crossval_folds(y = unique(imputations[[".id"]]), n_folds = n_folds)
-cross_val <- glmnet::cv.glmnet(
-    x = X_orig,
-    y = y_orig,
-    weights = wts_orig,
-    foldid = folds[match(imputations[[".id"]], folds[["y"]]), "fold"],
-    family = "binomial", 
-    alpha = 0
-  )
-
-plot(cross_val)
-
-dat_combined |> 
-  select(candidate_predictors, dataset) |>  
-    #Mc_FailedFemoralApproach, dataset) |> 
-  GGally::ggpairs(
-    aes(col = dataset, fill = dataset)
-  )
-
-  # First apparent performance - no penalty..
-mod <-glmnet::glmnet(
-    x = X_orig,
-    y = y_orig,
-    weights = wts_orig,
-    family = "binomial",
-    #lambda = 0, # unpenalized
-    lambda = cross_val$lambda.min,
-    alpha = 0, #ridge
-    intercept = TRUE
-  )
-
-hist(plogis(predict(mod, newx = X_orig)))
-max(plogis(predict(mod, newx = X_orig)))
-coef(mod)
-coef(
-  glm(model_formula, data = imps_combined, family = binomial())
-)
-calplot_MI(
-  imps_long = imps_combined,
-  model = mod,
-  impdat_ind = ".imp",
-  model_formula = model_formula,
-  height_hist = 0.2,
-  startpos_hist = -0.25,
-  knots = 5,
-  n_bins = 200,
-  xlim = c(0, 1),
-  ylim = c(-0.25, 1)
-) +
-  theme_minimal()
-
-# No penalty
-modo <- glm(
-  model_formula, family = binomial(), data = imps_combined, weights = imps_combined$wts
-)
-```
+(Ewout reference here, article)
 
@@
 
+### Internal validation combined cohort
+
+```{r combined-validation}
+imps_combined <- imps_all %>% 
+  filter(imps_label == "imps_combined")
+
+mod_combined <- validation_comb_lambdamin$model_fit
+validation_comb_lambdamin$model_fit$lambda
+mod_combined$beta
+
+validation_comb_lambdamin$validation_summary
+validation_comb_lambdamin$validation_df |> 
+  ggplot(aes(x = corrected)) +
+  geom_histogram(bins = 50) +
+  facet_wrap(~ measure, scales = "free")
+
+validation_comb_lambdamin$validation_df |> 
+  group_by(boot_num) |> 
+  summarise(pen = mean(penalty)) |>
+  #pull(pen) |>  mean()
+  ggplot(aes(pen)) +
+  geom_histogram()
+
+calplot_MI(
+  imps_long = imps_combined,
+  model = validation_comb_lambdamin$model_fit,
+  impdat_ind = ".imp",
+  model_formula = model_formula,
+  height_hist = 0.2,
+  startpos_hist = -0.25,
+  knots = 5,
+  n_bins = 200,
+  xlim = c(0, 0.55),
+  ylim = c(-0.25, 0.8)
+) +
+  theme_minimal()
+
+# Try simpler version of cal plot
+lp_combined <- predict(
+  validation_comb_lambdamin$model_fit,
+  newx = subset(
+    x = model.matrix(model_formula, data = imps_combined), 
+    select = -`(Intercept)`
+  )
+)
+
+df <- cbind.data.frame(
+      "predicted" = drop(lp_combined),
+      "outcome" = as.numeric(imps_combined$Mc_FailedFemoralApproach) - 1L,
+      "imp" = imps_combined$.imp
+    )
+
+binomial_smooth <- function(...) {
+  geom_smooth(method = "glm", method.args = list(family = "binomial"), ...)
+}
+
+ggplot(data = df, aes(plogis(predicted), outcome)) +
+  binomial_smooth(
+    aes(x = predicted),
+    formula = y ~ splines::ns(x, 3),
+    se = FALSE
+  ) #+
+  scale_x_continuous(trans = "log", limits = c(0.1, 1))
+  geom_smooth(
+    ,
+    aes(
+      x = predicted,
+      y = outcome#,
+      #group = imp
+    ), 
+    method = "glm",
+    #formula = y ~ splines::ns(x, 3),
+    method.args = list(family = "binomial"),
+    se = FALSE,
+    col = "gray",
+    alpha = 0.7
+  ) +
+  geom_abline(intercept = 0, slope = 1) +
+  lims(x = c(0, 1), y = c(0, 1))
+
+```
+
 ### Nomogram

### 2022-07-11 grp_resampling_efficiency.R
Sidetrack on efficient group resampling

@@ -1,0 +1,198 @@
+library(targets)
+library(tidyverse)
+
+tar_load(imps_all)
+
+imps_combined <- imps_all %>% 
+  filter(imps_label == "imps_combined")
+
+
+
+# All functions -----------------------------------------------------------
+
+
+# Original
+group_resample <- function(dat, 
+                           id_var, 
+                           B = 50, 
+                           boot_id_name = ".id_boot") {
+  
+  # Gather unique ids and resample
+  ids <- unique(dat[[id_var]])
+  
+  # Create bootstrap datasets
+  boot_dats <- replicate(
+    n = B,
+    simplify = FALSE,
+    expr = purrr::map_dfr(
+      .x = sample(x = ids, replace = TRUE), 
+      .f = ~ dat[which(dat[[id_var]] == .x), ], 
+      .id = boot_id_name
+    )
+  )
+  
+  return(boot_dats)
+}
+
+# For now all one bootstrap sample
+group_resample2 <- function(dat, 
+                            id_var, 
+                            B = 50, 
+                            boot_id_name = ".id_boot") {
+  
+  # Gather unique ids and resample
+  ids_long <- dat[[id_var]]
+  ids_unique <- unique(ids_long)
+  
+  # Get indicators
+  inds_long <- lapply(
+    X = sample(x = ids_unique, replace = TRUE), 
+    function(x) which(ids_long == x)
+  )
+  
+  dat[unlist(inds_long), ]
+}
+
+group_resample3 <- function(dat, 
+                            id_var, 
+                            B = 50, 
+                            boot_id_name = ".id_boot") {
+  
+  # Gather unique ids and resample
+  ids_long <- dat[[id_var]]
+  ids_unique <- unique(ids_long)
+  
+  # Sample, and prepare slices
+  samps <- sample(x = ids_unique, replace = TRUE)
+  inds_bigdat <- which(ids_long %in% samps)
+  slices_ls <- split(inds_bigdat, ids_long[inds_bigdat])
+  tbl <- table(samps)
+  slices_repped <- rep(slices_ls, times = tbl[match(names(slices_ls), names(tbl))])
+  
+  dat[unlist(slices_repped), ]
+}
+
+group_resample3_ext <- function(dat, 
+                                id_var, 
+                                B = 50, 
[…]

### 2022-07-16 analysis/article-results.Rmd
Removed logreg recal/started cleaning

@@
 library(tableone)
+library(rms)
+library(glmnet)
+```
 
+```{r objs}
 # Source functions
@@
 theme_set(theme_minimal(base_size = 14))
-```
 
-```{r objs}
 tar_load(
@@
 
-For continous variables:
+For continuous variables:
 
@@
 # Check proportion of complete cases in target analysis (quite good, compare results)
+
+# Development
 mean(complete.cases(dat_develop[, c("Mc_FailedFemoralApproach", candidate_predictors)]))
+
+# Validation
 mean(complete.cases(dat_valid[, c("Mc_FailedFemoralApproach", candidate_predictors)]))
+
+# Combined
 mean(complete.cases(dat_to_impute[, c("Mc_FailedFemoralApproach", candidate_predictors)]))
@@
 
-## Model update 1: logistic recalibration
-
-```{r log-recal}
-# 
-imps_valid$lp_valid <- lp_valid
-mod_recal <- glm(
-  Mc_FailedFemoralApproach ~ lp_valid, 
-  family = binomial(), 
-  data = imps_valid
-)
-boots_valid <- group_resample(imps_valid, id_var = ".id", 
-                              B = 50, 
-    boot_id_name = ".id_boot")
-
-boots_cals <- map_dfr(boots_valid, function(boot_dat) {
-  mod_recal_boot <- glm(
-  Mc_FailedFemoralApproach ~ lp_valid, 
-  family = binomial(), 
-  data = boot_dat
-)
-  calibration_intercept_slope(
-    imps_valid$Mc_FailedFemoralApproach, predict(mod_recal_boot, newdata = imps_valid)
-  )
-})
-colMeans(boots_cals)
-
-# Calib plot
-library(splines)
-ggplot() +
-  geom_smooth(
-    data = cbind.data.frame(
-      "predicted" = predict(mod_recal, type = "response"), #wrrooong
-      "outcome" = as.numeric(imps_valid$Mc_FailedFemoralApproach) - 1L,
-      "imp" = imps_valid$.imp
-    ),
-    aes(
-      x = predicted,
-      y = outcome,
-      group = imp
-    ), method = "glm",
-    formula = y ~ ns(x, 5),
-    method.args = list(family = "binomial"),
-    se = FALSE,
-    col = "gray",
-    alpha = 0.7
-  ) +
-  geom_abline(intercept = 0, slope = 1) +
-  lims(x = c(0, 1), y = c(0, 1))
-```
-
-## Model update 2: combined cohort re-estimation
+## Model update: combined cohort re-estimation
 
@@
 ```{r interaction-by-cohort}
+imps_combined <- imps_all %>% 
+  filter(imps_label == "imps_combined")
+
 split(imps_combined, ~ .imp) |> 
@@
 ```{r combined-validation}
-imps_combined <- imps_all %>% 
-  filter(imps_label == "imps_combined")
+mod_combined <- validation_comb_lambdamin$model_fit
+#validation_comb_lambdamin$model_fit$lambda
 
-mod_combined <- validation_comb_lambdamin$model_fit
-validation_comb_lambdamin$model_fit$lambda
-mod_combined$beta
+mod_coefs_comb <- as.matrix(coef(mod_combined))
+colnames(mod_coefs_comb) <- "Coefficients"
+
+# Print penalied model coefficients
+knitr::kable(data.frame(mod_coefs_comb))
 
@@
   theme_minimal()
-
-# Try simpler version of cal plot
-lp_combined <- predict(
-  validation_comb_lambdamin$model_fit,
-  newx = subset(
-    x = model.matrix(model_formula, data = imps_combined), 
-    select = -`(Intercept)`
-  )
-)
-
-df <- cbind.data.frame(
-      "predicted" = drop(lp_combined),
-      "outcome" = as.numeric(imps_combined$Mc_FailedFemoralApproach) - 1L,
-      "imp" = imps_combined$.imp
-    )
-
-binomial_smooth <- function(...) {
-  geom_smooth(method = "glm", method.args = list(family = "binomial"), ...)
-}
-
-ggplot(data = df, aes(plogis(predicted), outcome)) +
-  binomial_smooth(
-    aes(x = predicted),
-    formula = y ~ splines::ns(x, 3),
-    se = FALSE
-  ) #+
-  scale_x_continuous(trans = "log", limits = c(0.1, 1))
-  geom_smooth(
-    ,
-    aes(
-      x = predicted,
-      y = outcome#,
-      #group = imp
-    ), 
-    method = "glm",
-    #formula = y ~ splines::ns(x, 3),
-    method.args = list(family = "binomial"),
-    se = FALSE,
-    col = "gray",
-    alpha = 0.7
-  ) +
-  geom_abline(intercept = 0, slope = 1) +
-  lims(x = c(0, 1), y = c(0, 1))
-
 ```

### 2022-07-16 R/model-validation-helpers.R

@@ -28,6 +28,8 @@ group_resample <- function(dat,
   return(boot_dats)
 }
 
+# Above is not efficient
+
 imps_univariate_ORs <- function(outcome,
                                 predictors,
                                 imps,

### 2022-07-16 grp_resampling_efficiency.R

@@ -116,6 +116,65 @@ group_resample_dt <- function(dat,
   dplyr::slice(dat, slices_repped)
 }
 
+group_resample_dt2 <- function(dat, 
+                               id_var, 
+                               B = 50, 
+                               boot_id_name = ".id_boot") {
+  
+  # Gather unique ids and resample
+  ids_long <- dat[[id_var]]
+  
+  # Sample, and prepare slices
+  samps <- sample(x = unique(ids_long), replace = TRUE)
+  inds_bigdat <- which(ids_long %in% samps)
+  df <- data.table("ind" = inds_bigdat, "ids" = ids_long[inds_bigdat])
+  df[, reps := sum(samps == ids), by = ids]
+  slices_repped <- df[, .(inds_boot = rep(ind, reps))][["inds_boot"]]
+  dplyr::slice(dat, slices_repped)
+}
+
+
+group_resample4 <- function(dat, 
+                            id_var, 
+                            B = 50, 
+                            boot_id_name = ".id_boot") {
+  
+  # Gather unique ids and resample
+  setDT(dat)
+  ids_long <- dat[[id_var]]
+  
+  # Sample, and prepare slices (not as.numeric if making more general)
+  boot_dats <- lapply(seq_len(B), function(b) {
+    #cat(b)
+    samps <- sample(x = unique(ids_long), replace = TRUE)
+    inds_bigdat <- which(ids_long %in% samps)
+    ids_bigdat <- ids_long[inds_bigdat]
+    tbl <- tabulate(samps, nbins = length(samps))
+    dat[rep(inds_bigdat, times = tbl[ids_bigdat])]
+  })
+  
+  return(boot_dats)
+}
+
+
+testo <- group_resample4(
+  dat = imps_combined,
+  id_var = ".id",
+  B = 200
+)
+
+res <- microbenchmark::microbenchmark(
+  group_resample4(
+    dat = imps_combined,
+    id_var = ".id",
+    B = 20
+  ),
+  times = 50L
+)
+boxplot(res)
+ggplot2::autoplot(res)
+
+
 #
 microbenchmark::microbenchmark(
   "orig" = {
@@ -150,20 +209,20 @@ microbenchmark::microbenchmark(
       boot_id_name = ".id_boot"
     )
   },
-  times = 5L
-)
-
-
-# The beasts
-microbenchmark::microbenchmark(
-  "slices_rep" = {
-    group_resample3(
[…]

### 2022-07-26 analysis/article-results.Rmd
Start of cleanup for proper interaction-by-cohort

@@
 ---
-title: "Article results"
+title: "Validation failed EVT"
 author: "E. F. Bonneville"
@@
 library(glmnet)
+library(DT)
 ```
@@
 # Summaries of all continuous variables
-kableone(table_one$ContTable, nonnormal = TRUE)
+invisible(
+  capture.output(
+    tab_contin <- print(
+      x = table_one$ContTable, 
+      nonnormal = TRUE, 
+      noSpaces = TRUE
+    )
+  )
+)
+
+#https://stackoverflow.com/questions/50039186/add-download-buttons-in-dtrenderdatatable
+DT::datatable(
+  tab_contin,
+  extensions = 'Buttons',
+  options = list(
+    dom = 'Blfrtip', 
+    buttons = c('csv', 'excel', 'pdf'),
+    lengthMenu = list(c(10, 25, 50, -1), c(10, 25, 50, "All"))
+  )
+)
 ```
@@
 # Summaries of all categorical variables
-kableone(table_one$CatTable)
+invisible(
+  capture.output(
+    tab_cat <- print(
+      x = table_one$CatTable, 
+      nonnormal = TRUE, 
+      noSpaces = TRUE
+    )
+  )
+)
+
+DT::datatable(
+  tab_cat,
+  extensions = 'Buttons',
+  options = list(
+    dom = 'Blfrtip', 
+    buttons = c('csv', 'excel', 'pdf'),
+    lengthMenu = list(c(10, 25, 50, -1), c(10, 25, 50, "All"))
+  )
+)
 ```
@@
 
-```{r miss-dat-summ}
+```{r cca-check}
 # Seperate datasets
 dat_develop <- subset(x = dat_to_impute, select = -dataset, subset = (dataset == "develop"))
+
 dat_valid <- subset(x = dat_to_impute, select = -dataset, subset = (dataset == "valid"))
 
-# Check proportion of complete cases in target analysis (quite good, compare results)
-
 # Development
-mean(complete.cases(dat_develop[, c("Mc_FailedFemoralApproach", candidate_predictors)]))
+prop_cca_dev <- mean(complete.cases(dat_develop[, c("Mc_FailedFemoralApproach", candidate_predictors)]))
 
 # Validation
-mean(complete.cases(dat_valid[, c("Mc_FailedFemoralApproach", candidate_predictors)]))
+prop_cca_valid <- mean(complete.cases(dat_valid[, c("Mc_FailedFemoralApproach", candidate_predictors)]))
 
 # Combined
-mean(complete.cases(dat_to_impute[, c("Mc_FailedFemoralApproach", candidate_predictors)]))
+prop_cca_comb <- mean(complete.cases(dat_to_impute[, c("Mc_FailedFemoralApproach", candidate_predictors)]))
+```
 
+Complete-case analysis (using predictors in the prediction model) results in:
+
+- `r paste0(round(prop_cca_dev * 100, 2), " %")` complete cases in the development data
+- `r paste0(round(prop_cca_valid * 100, 2), " %")` complete cases in the validation data
+- `r paste0(round(prop_cca_comb * 100, 2), " %")` complete cases in the combined data
+
+```{r miss-dat-summ}
 # Visualise missings beforehand
@@
 
-# Print penalied model coefficients
-knitr::kable(data.frame(mod_coefs))
+# Print penalised model coefficients
+#knitr::kable(data.frame(mod_coefs))
+DT::datatable(
+  data.frame(round(mod_coefs, digits = 3)),
+  extensions = 'Buttons',
+  options = list(
+    dom = 'Blfrtip', 
+    buttons = c('csv', 'excel', 'pdf'),
+    lengthMenu = list(c(10, 25, 50, -1), c(10, 25, 50, "All"))
+  )
+)
 ```
@@
 
-psfmi::pool_auc(auc_df$auc, sqrt(auc_df$var_auc), nimp = max(as.numeric(auc_df$.imp)))
+auc_ext <- psfmi::pool_auc(
+  est_auc = auc_df$auc,
+  est_se = sqrt(auc_df$var_auc),
+  nimp = max(as.numeric(auc_df$.imp))
+)
 
 # Get cal slope and intercept
-split(df_calplot_valid, ~ .imp) |> 
+cal_slope_ext <- split(df_calplot_valid, ~ .imp) |> 
   map(.f = ~ glm(Mc_FailedFemoralApproach ~ lp_valid, family = binomial, data = .x)) |> 
   mice::pool() |> 
-  summary()
+  summary(conf.int = TRUE)
 
-split(df_calplot_valid, ~ .imp) |> 
+cal_int_ext <- split(df_calplot_valid, ~ .imp) |> 
   map(.f = ~ glm(Mc_FailedFemoralApproach ~ offset(lp_valid), family = binomial, data = .x)) |> 
   mice::pool() |> 
-  summary()
+  summary(conf.int = TRUE)
+
+int_ext <- with(
+  cal_int_ext, 
+  paste0(
+    round(estimate, 3), " [",
+    round(`2.5 %`, 3), ", ",
+    round(`97.5 %`, 3), "]"
+  )
+)
+
+slope_ext <- with(
+  cal_slope_ext[2, ], 
+  paste0(
+    round(estimate, 3), " [",
+    round(`2.5 %`, 3), ", ",
+    round(`97.5 %`, 3), "]"
+  )
+)
+
+auc_ext_bis <- with(
+  as.data.frame(auc_ext),
+  paste0(
+    round(`C-statistic`, 3), " [",
+    round(`95% Low`, 3), ", ",
+    round(`95% Up`, 3), "]"
+  )
+)
+
+rbind.data.frame(
+  cbind.data.frame("measure" = "auc", "value" = auc_ext_bis),
+  cbind.data.frame("measure" = "intercept", "value" = int_ext),
+  cbind.data.frame("measure" = "slope", "value" = slope_ext)
+)
 

### 2022-07-26 analysis/article-results.Rmd
First versio with nomogram

@@
 
-split(imps_combined, ~ .imp) |> 
+inter_model <- split(imps_combined, ~ .imp) |> 
   map(.f = ~ glm(update(model_formula, . ~ . * dataset), family = binomial, data = .x)) |> 
   mice::pool() |> 
-  broom::tidy() |>  
-  ggforestplot::forestplot(
-    name = term,
-    estimate = estimate,
-    se = std.error
+  broom::tidy()
+
+summ_inter <- cbind.data.frame(
+  inter_model |>  
+    filter(str_detect(term, pattern = "valid$", negate = TRUE)) |> 
+    select(term),
+  "development" = inter_model |>  
+    filter(str_detect(term, pattern = "valid$", negate = TRUE)) |> 
+    pull(estimate),
+  "interaction" = inter_model |>  
+    filter(str_detect(term, pattern = "valid$")) |> 
+    pull(estimate),
+  "pval_interaction" = inter_model |>  
+    filter(str_detect(term, pattern = "valid$")) |> 
+    pull(p.value)
+) |> 
+  mutate(
+    validation = development + interaction
+  ) |> 
+  mutate(across(where(is.numeric), ~ round(.x, 3)))
+
+DT::datatable(
+  summ_inter,
+  extensions = 'Buttons',
+  options = list(
+    dom = 'Blfrtip', 
+    buttons = c('csv', 'excel', 'pdf'),
+    lengthMenu = list(c(10, 25, 50, -1), c(10, 25, 50, "All"))
   )
-
-split(imps_combined, ~ .imp) |> 
-  map(.f = ~ glm(update(model_formula, . ~ . * dataset), family = binomial, data = .x)) |> 
-  mice::pool() |> 
-  broom::tidy() |> 
-  select(term, estimate)
-
-split(imps_combined |> filter(dataset == "develop"), ~ .imp) |> 
-  map(.f = ~ glm(model_formula, family = binomial, data = .x)) |> 
-  mice::pool() |> 
-  broom::tidy() |> 
-  pull(estimate)
-
-cbind.data.frame(
-  split(imps_combined, ~ .imp) |> 
-  map(.f = ~ glm(update(model_formula, . ~ . * dataset), family = binomial, data = .x)) |> 
-  mice::pool() |> 
-  broom::tidy() |> 
-  select(term, estimate),
-  "develop" = split(imps_combined |> filter(dataset == "develop"), ~ .imp) |> 
-  map(.f = ~ glm(model_formula, family = binomial, data = .x)) |> 
-  mice::pool() |> 
-  broom::tidy() |> 
-  pull(estimate),
-  "valid" =  split(imps_combined |> filter(dataset == "valid"), ~ .imp) |> 
-  map(.f = ~ glm(model_formula, family = binomial, data = .x)) |> 
-  mice::pool() |> 
-  broom::tidy() |> 
-  pull(estimate),
-  "combined" =  split(imps_combined, ~ .imp) |> 
-  map(.f = ~ glm(model_formula, family = binomial, data = .x)) |> 
-  mice::pool() |> 
-  broom::tidy() |> 
-  pull(estimate)
-) |> 
-  View()
-
-# ORs
-cbind.data.frame(
-  split(imps_combined, ~ .imp) |> 
-  map(.f = ~ glm(update(model_formula, . ~ . * dataset), family = binomial, data = .x)) |> 
-  mice::pool() |> 
-  broom::tidy(exponentiate = TRUE) |> 
-  select(term, estimate),
-  "develop" = split(imps_combined |> filter(dataset == "develop"), ~ .imp) |> 
-  map(.f = ~ glm(model_formula, family = binomial, data = .x)) |> 
-  mice::pool() |> 
-  broom::tidy(exponentiate = TRUE) |> 
-  pull(estimate),
-  "valid" =  split(imps_combined |> filter(dataset == "valid"), ~ .imp) |> 
-  map(.f = ~ glm(model_formula, family = binomial, data = .x)) |> 
-  mice::pool() |> 
-  broom::tidy(exponentiate = TRUE) |> 
-  pull(estimate),
-  "combined" =  split(imps_combined, ~ .imp) |> 
-  map(.f = ~ glm(model_formula, family = binomial, data = .x)) |> 
-  mice::pool() |> 
-  broom::tidy(exponentiate = TRUE) |> 
-  pull(estimate)
-) |> 
-  View()
-
-model.matrix(model_formula, data=imps_combined) %>% 
-  cor(use="pairwise.complete.obs") %>% 
-  ggcorrplot::ggcorrplot(show.diag = F, type="lower", lab=TRUE, lab_size=2)
+)
 ```
@@
 
-# Print penalied model coefficients
-knitr::kable(data.frame(mod_coefs_comb))
+# Print penalised model coefficients
+DT::datatable(
+  data.frame(round(mod_coefs_comb, digits = 3)),
+  extensions = 'Buttons',
+  options = list(
+    dom = 'Blfrtip', 
+    buttons = c('csv', 'excel', 'pdf'),
+    lengthMenu = list(c(10, 25, 50, -1), c(10, 25, 50, "All"))
+  )
+)
 
-validation_comb_lambdamin$validation_summary
-validation_comb_lambdamin$validation_df |> 
-  ggplot(aes(x = corrected)) +
-  geom_histogram(bins = 50) +
-  facet_wrap(~ measure, scales = "free")
-
-validation_comb_lambdamin$validation_df |> 
-  group_by(boot_num) |> 
-  summarise(pen = mean(penalty)) |>
-  #pull(pen) |>  mean()
-  ggplot(aes(pen)) +
-  geom_histogram()
+# Optimism corrected
+validation_comb_lambdamin$validation_summary %>% 
+  filter(measure %in% c("auc", "intercept", "slope")) %>% 
+  group_by(measure) %>% 
+  summarise(value = paste0(corrected, " [", lower_corrected, ";", upper_corrected, "]")) %>%
+  knitr::kable()
 
@@
 ### Nomogram
+
+```{r nomogram}
+# Make data for nomogram labels
+dat_nomogram <- imps_combined %>% 
+  select(all_of(c(response_var, candidate_predictors))) %>% 
+  mutate(
+    InnCca_nr_90orLarger = factor(
+      InnCca_nr_90orLarger,
+      levels = levels(InnCca_nr_90orLarger),
+      labels = c("0", ">=1")
+    ),
+    ICAE_NASCET_Degree = factor(
+      ICAE_NASCET_Degree,
+      levels = levels(ICAE_NASCET_Degree),
+      labels = c("<99", ">=99")
+    )
+  )
+
+# Set variable labels
+Hmisc::label(dat_nomogram$M_age) <- "Age (decades)"
+Hmisc::label(dat_nomogram$M_prev_ht) <- "Hypertension"
+Hmisc::label(dat_nomogram$AorticVariant) <- "Aortic variant"
+Hmisc::label(dat_nomogram$ArchElongation) <- "Arch Elongation"
+Hmisc::label(dat_nomogram$AngleFollowingBifurcation) <- "Angle post-bifurcation"
+Hmisc::label(dat_nomogram$InnCca_nr_90orLarger) <- "CCA angles >=90 degr."
+Hmisc::label(dat_nomogram$ICAE_NASCET_Degree) <- "ICA stenosis"
+
+dd <- datadist(dat_nomogram)
+options(datadist = "dd")
+
+# Edit some labels
+mod_lrm <- lrm(model_formula, data = dat_nomogram)
+mod_lrm$coefficients <- setNames(unname(drop(mod_coefs_comb)), names(mod_lrm$coefficients))
+
+par(mar = c(2,2,1,1))
+nom <- nomogram(
+  fit = mod_lrm, 
+  fun = plogis, 
+  maxscale = 10, 
+  funlabel = "Predicted probability", 
+  lp = FALSE, 
+  fun.at = c(.01, .05, seq(.1, 1, by = .1))
+)
+
+plot(nom)
+
+# Get max probability in combined dataset?
+```

### 2022-07-28 analysis/article-results.Rmd
Rmd including figure exports

@@
 
+The next plots look at the combinations of variables missings and their frequencies in the development and validation sets respectively:
+
 ```{r miss-dat-summ}
@@
 naniar::gg_miss_var(dat_to_impute, facet = dataset, show_pct = TRUE)
+naniar::gg_miss_upset(dat_develop, nsets = 10, )
 naniar::gg_miss_upset(dat_valid, nsets = 10)
-naniar::gg_miss_upset(dat_develop, nsets = 10)
 ```
@@
 ## Internal validation development set
+
+Coefficients of the (penalized, with Ridge) prediction model built using the development data:
 
@@
 
-Bootstrap-based internal validation, optimism-corrected performance:
+Bootstrap-based internal validation, optimism-corrected performance of this model:
 
@@
 
-Calibration plot (apparent)
+Calibration plot, reflecting apparent performance:
 
@@
 
+At external validation, the calibration plot shows that there is quite some overestimation of risks in the validation set.
+
 ## Model update: combined cohort re-estimation
@@
 
-(Ewout reference here, article)
+See [Balmaña, Stockwell, Steyerberg et al.
+(2006)](https://jamanetwork.com/journals/jama/article-abstract/203427).
+Here, we use standard (unpenalized) logistics regression to test for differences between the development and validation set in terms of multivariable model coefficients.
+
+The log odds table of coefficient, and the p-values for the interaction:
 
@@
   summ_inter,
+  extensions = 'Buttons',
+  options = list(
+    dom = 'Blfrtip', 
+    buttons = c('csv', 'excel', 'pdf'),
+    lengthMenu = list(c(10, 25, 50, -1), c(10, 25, 50, "All"))
+  )
+)
+```
+
+In the format of Table 4 from the article mentioned above (with odds ratios):
+
+```{r inter-steyerberg}
+# Second attempt in Ewout format 
+# 
+library(broom.helpers)
+library(broom)
+
+dev_unpen <- split(imps_combined |> 
+                     filter(dataset == "develop"), ~ .imp) |> 
+  map(.f = ~ glm(model_formula, family = binomial, data = .x)) |> 
+  mice::pool() |> 
+  #broom::tidy(exponentiate = TRUE, conf.int = TRUE) |> 
+  tidy(exponentiate = TRUE, conf.int = TRUE) |> 
+  #tidy_add_reference_rows(no_reference_row = all_categorical())
+  mutate(
+    "Development" = paste0(
+      round(estimate, 3), " [",
+      round(conf.low, 3), ", ",
+      round(conf.high, 3), "]"
+    )
+  ) |> 
+  select(term, Development)
+
+valid_unpen <- split(imps_combined |> 
+                     filter(dataset == "valid"), ~ .imp) |> 
+  map(.f = ~ glm(model_formula, family = binomial, data = .x)) |> 
+  mice::pool() |> 
+  #broom::tidy(exponentiate = TRUE, conf.int = TRUE) |> 
+  tidy(exponentiate = TRUE, conf.int = TRUE) |> 
+  #tidy_add_reference_rows(no_reference_row = all_categorical())
+  mutate(
+    "Validation" = paste0(
+      round(estimate, 3), " [",
+      round(conf.low, 3), ", ",
+      round(conf.high, 3), "]"
+    )
+  ) |> 
+  select(term, Validation)
+
+comb_unpen <- split(imps_combined, ~ .imp) |> 
+  map(.f = ~ glm(model_formula, family = binomial, data = .x)) |> 
+  mice::pool() |> 
+  #broom::tidy(exponentiate = TRUE, conf.int = TRUE) |> 
+  tidy(exponentiate = TRUE, conf.int = TRUE) |> 
+  #tidy_add_reference_rows(no_reference_row = all_categorical())
+  mutate(
+    "Combined" = paste0(
+      round(estimate, 3), " [",
+      round(conf.low, 3), ", ",
+      round(conf.high, 3), "]"
+    )
+  ) |> 
+  select(term, Combined)
+
+cbind.data.frame(
+  "Term" = dev_unpen$term, 
+  "Development \n cohort" = dev_unpen$Development, 
+  "Validation \n cohort" = valid_unpen$Validation,
+  "Combined \n cohort" = comb_unpen$Combined
+) |> 
+  DT::datatable(
   extensions = 'Buttons',
@@
 ```
+
+```{r nom-export, results='hide'}
+png(
+  filename = "analysis/figures/nomogram-combined.png",
+  width = 8,
+  height = 11,
+  units = "in",
+  res = 300
+)
+plot(nom)
+dev.off()
+
+# Predict max prob
+newpat <- dat_nomogram[1, ]
+newpat$ArchElongation <- "III"
+newpat$AngleFollowingBifurcation <- "Yes, angle >=90 degrees"
+newpat$InnCca_nr_90orLarger <- ">=1"
+newpat$ICAE_NASCET_Degree <- ">=99"
+newpat$M_age <- 8
+newpat
+
+predict(mod_lrm, type = "fitted", newdata = newpat)
+```

### 2022-07-28 analysis/validate-models.R

@@ -1,547 +1,0 @@
-##**************************##
-## Post-imputation analyses ##
-##**************************##
-
-
-# Run from start to finish - I have commented out code that should
-# not be re-run.
-
-# Set contrasts for ordered factors
-options(contrasts = rep("contr.treatment", 2))
-
-# Read-in custom functions for validation
-source("analysis/helper-functions.R")
-
-# To reproduce all bootstraps
-set.seed(1202)
-
-candidate_predictors <- c(
-  "M_age",
-  "M_prev_ht",
-  "AorticVariant",
-  "ArchElongation",
-  "AngleFollowingBifurcation",
-  "InnCca_nr_90orLarger",
-  "ICAE_NASCET_Degree"
-)
-
-# General model formula - keep response character
-form <- reformulate(termlabels = candidate_predictors, response = "Mc_FailedFemoralApproach")
-response_var <- all.vars(stats::update(form, . ~ 1))
-form
-
-# Prepare variables  ------------------------------------------------------
-
-
-# Prepare to read-in imputations
-imps_files <- list.files(path = "data/", pattern = "^imps", full.names = TRUE)
-names(imps_files) <- gsub(x = imps_files, pattern = "^data/|\\.rds$", replacement = "")
-
-# Read all in together to transform variables at once (Julia's code)
-imps_all <- imps_files %>% 
-  map(.f = ~ mice::complete(readRDS(.x), action = "long")) %>% 
-  bind_rows(.id = "imps_label") %>% 
-  mutate(
-    ICAE_NASCET_Degree = factor(
-      x = ifelse(ICAE_NASCET_Degree < 99, "leq_99", "geq_99"), 
-      levels = c("leq_99", "geq_99")
-    ),
-    InnCca_nr_90orLarger = fct_collapse(
-      .f = InnCca_nr_90orLarger, 
-      "0" = "0", 
-      "geq_1" = c("1", "2", "3")
-    ),
-    ICA_nr_90orLarger = fct_collapse(
-      .f = ICA_nr_90orLarger, 
-      "0" = "0", 
-      "geq_1" = c("1", "2", "3")
-    ),
-    ICAIAtherosclerosis = fct_collapse(
-      .f = ICAIAtherosclerosis, 
-      "under50%" = c("No", "Yes, calcified spots", "Yes, <50% stenosis"),
-      "geq50%" = "Yes, >=50% stenosis"
-    ),
[…]

### 2022-09-26 analysis/article-results.Rmd
Updates to table 1/nomogram

@@
 ```{r table-one-contin}
+# ?tar_mermaid()
+
 # Gives you everything for tables 1 and 2 (warnings due to variables not in validation, ignore!)
-table_one <- dat_combined %>%  
-  select(-all_of(c("sidescored", "AngleCcaIca", "ICAEAtherosclerosis", "StudySubjectID"))) %>%
+table_one <- select(.data = tar_read(develop_n887_raw), c("StudySubjectID", "M_mrs_rev")) |> 
+  bind_rows(select(.data = tar_read(valid_n1111_raw), c("StudySubjectID", "M_mrs_rev"))) |> 
+  rename("M_mrs_rev_orig" = "M_mrs_rev") |> 
+  left_join(dat_combined) %>%  
+  select(-all_of(c("sidescored", "AngleCcaIca", "ICAEAtherosclerosis", "StudySubjectID", "M_mrs_rev"))) %>%
   # Here we dichotomise purely for the table, does not affect original data
@@
     ICAI_stenosis50 = factor(ifelse(ICAIAtherosclerosis == "Yes, >=50% stenosis", 1, 0)),
-    AngleAaInn_OR_AaCca_dich45 = factor(ifelse(AngleAaInn_OR_AaCca > 45, 0, 1))
+    AngleAaInn_OR_AaCca_dich45 = factor(ifelse(AngleAaInn_OR_AaCca > 45, 0, 1)),
+    Mc_pretici_short = fct_collapse(
+      Mc_pretici_short, 
+      "<1" = c("0"),
+      ">=1" = c("1", "2A", "2B, 2C or 3")
+    ),
+    M_collaterals = fct_collapse(
+      M_collaterals,
+      "<50" = c("absent collaterals", "filling <50% of occluded area"),
+      ">=50" = c(">50% but less <100%", "100% of occluded area")
+    ),
+    M_ASPECTS_BL = factor(
+      ifelse(M_ASPECTS_BL > 7, 0, 1),
+      levels = c(0, 1),
+      labels = c("8-10", "<=7")
+    ),
+    M_CBS_BL = factor(
+      ifelse(M_CBS_BL > 7, 0, 1),
+      levels = c(0, 1),
+      labels = c("8-10", "<=7")
+    ),
+    M_posttici_c = fct_collapse(
+      M_posttici_c,
+      "2B-3" = c("2B", "2C", "3"),
+      "0-2A" = c("0", "1", "2A")
+    ),
+    M_mrs_rev_orig = factor(
+      ifelse(M_mrs_rev_orig >= 4, 0, 1),
+      levels = c(0, 1),
+      labels = c("0-2", "3-6")
+    )
   ) %>% 
@@
 naniar::gg_miss_var(dat_to_impute, facet = dataset, show_pct = TRUE)
-naniar::gg_miss_upset(dat_develop, nsets = 10, )
-naniar::gg_miss_upset(dat_valid, nsets = 10)
+#naniar::gg_miss_upset(dat_develop, nsets = 10, )
+#naniar::gg_miss_upset(dat_valid, nsets = 10)
+
+dat_to_impute |> 
+  group_by(dataset) |> 
+  naniar::miss_var_summary() |> 
+  mutate(across(where(is.numeric), ~ round(.x, 2L))) |> 
+  DT::datatable(
+    extensions = 'Buttons',
+    options = list(
+      dom = 'Blfrtip', 
+      buttons = c('csv', 'excel', 'pdf'),
+      lengthMenu = list(c(10, 25, 50, -1), c(10, 25, 50, "All"))
+    )
+  )  
 ```
@@
 # Overestimation of risks
-calplot_MI(
+calplot_ext <- calplot_MI(
   imps_long = imps_valid,
@@
   n_bins = 200,
-  xlim = c(0, 0.85),
+  xlim = c(0, 0.7),
   ylim = c(-0.25, 1)
 ) +
-  theme_minimal()
+  theme_minimal() + 
+  labs(x = "Predicted probability", y = "Observed proportion")
+
+calplot_ext
+
+ggsave(
+  calplot_ext,
+  filename = "analysis/figures/calibration-plot-ext.png",
+  width = 8,
+  height = 5,
+  units = "in",
+  dpi = 300, 
+  bg = "white"
+)
 ```
@@
 
-calplot_MI(
+calplot_combined <- calplot_MI(
   imps_long = imps_combined,
@@
 ) +
-  theme_minimal()
+  theme_minimal() +
+  labs(x = "Predicted probability", y = "Observed proportion")
+
+calplot_combined
+
+ggsave(
+  calplot_combined,
+  filename = "analysis/figures/calibration-plot-combined.png",
+  width = 8,
+  height = 5,
+  units = "in",
+  dpi = 300, bg = "white"
+)
 ```
@@
 # Set variable labels
-Hmisc::label(dat_nomogram$M_age) <- "Age (decades)"
+Hmisc::label(dat_nomogram$M_age) <- "Age (per decade)"
 Hmisc::label(dat_nomogram$M_prev_ht) <- "Hypertension"
-Hmisc::label(dat_nomogram$AorticVariant) <- "Aortic variant"
-Hmisc::label(dat_nomogram$ArchElongation) <- "Arch Elongation"
-Hmisc::label(dat_nomogram$AngleFollowingBifurcation) <- "Angle post-bifurcation"
-Hmisc::label(dat_nomogram$InnCca_nr_90orLarger) <- "CCA angles >=90 degr."
+Hmisc::label(dat_nomogram$ArchElongation) <- "Aortic Arch Elongation"
+Hmisc::label(dat_nomogram$AorticVariant) <- "Aortic variants"
+Hmisc::label(dat_nomogram$InnCca_nr_90orLarger) <- "Tortuosity IA and CCA"
+Hmisc::label(dat_nomogram$AngleFollowingBifurcation) <- "Tortuosity cervical ICA"
 Hmisc::label(dat_nomogram$ICAE_NASCET_Degree) <- "ICA stenosis"
@@
 # Edit some labels
+model_formula_reorder <- Mc_FailedFemoralApproach ~ M_age + M_prev_ht + ArchElongation +
+  AorticVariant + InnCca_nr_90orLarger + AngleFollowingBifurcation + ICAE_NASCET_Degree
+
+# Annoying stuff to reorder
+mod_lrm_reorder <- lrm(model_formula_reorder, data = dat_nomogram)
 mod_lrm <- lrm(model_formula, data = dat_nomogram)
 mod_lrm$coefficients <- setNames(unname(drop(mod_coefs_comb)), names(mod_lrm$coefficients))
+mod_lrm_reorder$coefficients <- mod_lrm$coefficients[match(
+  names(mod_lrm_reorder$coefficients), names(mod_lrm$coefficients)
+)]
 
@@
 nom <- nomogram(
-  fit = mod_lrm, 
+  fit = mod_lrm_reorder, 
   fun = plogis, 
@@
 
+# Adjust levels
+nom$`Tortuosity cervical ICA`$AngleFollowingBifurcation <- c("<90°", ">=90°") #°
+nom$`Tortuosity IA and CCA`$InnCca_nr_90orLarger <- c("<90°", ">=90°") #°
+nom$`ICA stenosis`$ICAE_NASCET_Degree <- c("<99%", ">=99%") #°
+nom$`Aortic variants`$AorticVariant <- c("Normal", "Common origin\nIA-CCA", "Bovine arch") #°
+nom$`Aortic Arch Elongation`$ArchElongation <- c("No elongation", "Mild to moderate", "Severe") #°
+
 plot(nom)
+
+# Printing points
+print(nom, dec = 2L)
 
@@
 newpat$ICAE_NASCET_Degree <- ">=99"
-newpat$M_age <- 8
+newpat$M_age <- 10
 newpat
@@
 predict(mod_lrm, type = "fitted", newdata = newpat)
+dat_nomogram[which.max(predict(mod_lrm, type = "fitted", newdata = dat_nomogram)), ]
+predict(mod_lrm, type = "fitted", newdata = dat_nomogram) |>  max()
 ```

### 2022-09-27 grp_resampling_efficiency.R
Further cleaning of files

@@ -1,264 +1,0 @@
-library(targets)
-library(tidyverse)
-
-tar_load(imps_all)
-
-imps_combined <- imps_all %>% 
-  filter(imps_label == "imps_combined")
-
-
-
-# All functions -----------------------------------------------------------
-
-
-# Original
-group_resample <- function(dat, 
-                           id_var, 
-                           B = 50, 
-                           boot_id_name = ".id_boot") {
-  
-  # Gather unique ids and resample
-  ids <- unique(dat[[id_var]])
-  
-  # Create bootstrap datasets
-  boot_dats <- replicate(
-    n = B,
-    simplify = FALSE,
-    expr = purrr::map_dfr(
-      .x = sample(x = ids, replace = TRUE), 
-      .f = ~ dat[which(dat[[id_var]] == .x), ], 
-      .id = boot_id_name
-    )
-  )
-  
-  return(boot_dats)
-}
-
-# For now all one bootstrap sample
-group_resample2 <- function(dat, 
-                            id_var, 
-                            B = 50, 
-                            boot_id_name = ".id_boot") {
-  
-  # Gather unique ids and resample
-  ids_long <- dat[[id_var]]
-  ids_unique <- unique(ids_long)
-  
-  # Get indicators
-  inds_long <- lapply(
-    X = sample(x = ids_unique, replace = TRUE), 
-    function(x) which(ids_long == x)
-  )
-  
-  dat[unlist(inds_long), ]
-}
-
-group_resample3 <- function(dat, 
-                            id_var, 
-                            B = 50, 
-                            boot_id_name = ".id_boot") {
-  
-  # Gather unique ids and resample
-  ids_long <- dat[[id_var]]
-  ids_unique <- unique(ids_long)
-  
-  # Sample, and prepare slices
-  samps <- sample(x = ids_unique, replace = TRUE)
-  inds_bigdat <- which(ids_long %in% samps)
-  slices_ls <- split(inds_bigdat, ids_long[inds_bigdat])
-  tbl <- table(samps)
-  slices_repped <- rep(slices_ls, times = tbl[match(names(slices_ls), names(tbl))])
-  
-  dat[unlist(slices_repped), ]
-}
-
-group_resample3_ext <- function(dat, 
-                                id_var, 
-                                B = 50, 
[…]

### 2022-09-27 run-pipeline.R

@@ -1,8 +1,0 @@
-# Visualise
-tar_visnetwork()
-
-# Run/reproduce full analysis
-#tar_make()
-tar_make_future(workers = availableCores() - 1) # use min(1, ...) formulation
-
-#tar_meta(fields = warnings)

### 2023-03-30 analysis/article-results.Rmd
Testing md addition

@@
   html_document:
+    keep_md: true
     df_print: kable
@@
 
-# Load libraries for knitting
+# Load libraries just for analysis rmd
 library(ggplot2)
@@
 library(DT)
+library(broom.helpers)
+library(broom)
 ```
@@
 # Second attempt in Ewout format 
-# 
-library(broom.helpers)
-library(broom)
-
 dev_unpen <- split(imps_combined |> 

### 2023-03-30 analysis/article-results.Rmd
rmd is part of targets pipeline, and 2nd test md

@@
   chunk_output_type: console
+params:
+  interactive_tables: FALSE
 ---
@@
 library(broom)
+library(knitr)
 ```
@@
 #https://stackoverflow.com/questions/50039186/add-download-buttons-in-dtrenderdatatable
+if (params$interactive_tables) {
   DT::datatable(
@@
   )
+} else kableone(table_one$ContTable)
+
 ```
@@
 
+if (params$interactive_tables) {
   DT::datatable(
@@
   )
+} else kableone(table_one$CatTable)
 ```
@@
 
-dat_to_impute |> 
+df_missings <- dat_to_impute |> 
   group_by(dataset) |> 
   naniar::miss_var_summary() |> 
-  mutate(across(where(is.numeric), ~ round(.x, 2L))) |> 
+  mutate(across(where(is.numeric), ~ round(.x, 2L)))
+
+if (params$interactive_tables)  {
   DT::datatable(
+    df_missings,
     extensions = 'Buttons',
@@
   )  
+} else kable(df_missings)
 ```
@@
 #knitr::kable(data.frame(mod_coefs))
+
+if (params$interactive_tables)  {
+
   DT::datatable(
@@
   )
+} else kable(data.frame(round(mod_coefs, digits = 3)))
 ```
@@
   )
+} else kable(summ_inter)
 ```
@@
 
-cbind.data.frame(
+df_unpen <- cbind.data.frame(
   "Term" = dev_unpen$term, 
-  "Development \n cohort" = dev_unpen$Development, 
-  "Validation \n cohort" = valid_unpen$Validation,
-  "Combined \n cohort" = comb_unpen$Combined
-) |> 
+  "Development cohort" = dev_unpen$Development, 
+  "Validation cohort" = valid_unpen$Validation,
+  "Combined cohort" = comb_unpen$Combined
+)
+
+if (params$interactive_tables) {
   DT::datatable(
+    df_unpen,
     extensions = 'Buttons',
@@
   )
+} else kable(df_unpen)
 ```
@@
 # Print penalised model coefficients
+if (params$interactive_tables) {
   DT::datatable(
@@
   )
+} else kable(data.frame(round(mod_coefs_comb, digits = 3)))
 
@@
 
+dd <<- datadist(dat_nomogram)
+options(datadist = "dd")
+
 # Set variable labels
@@
 Hmisc::label(dat_nomogram$ICAE_NASCET_Degree) <- "ICA stenosis"
-
-dd <- datadist(dat_nomogram)
-options(datadist = "dd")
 

### 2023-03-30 _targets.R

@@ -42,8 +42,8 @@ source("data-raw/prepare_raw_data.R")
 source("R/imputation-helpers.R")
 source("R/model-validation-helpers.R")
 
-# Pipeline (parts are in separate files):
-targets_list <- list(
+# Pipeline 
+list(
 
   # -- Part 1: Data preparation
   tar_target(
@@ -133,16 +133,8 @@ targets_list <- list(
       lambda_choice = "min",
       B = validation_settings$B
     )
-  )
+  ),
+  tarchetypes::tar_render(analysis_summary, path = "analysis/article-results.Rmd")
   # 
   #tarchetypes::tar_render(analysis_summary, path = "analysis/2020-09_analysis-summary.Rmd")
 )
-
-# Source targets
-#("R/NMA-preDLI-models.R")
-#source("R/NMA-postDLI-models.R")
-
-#targets_list <- c(targets_list, bind_imps)
-targets_list
-
-    

### 2023-03-30 analysis/article-results.Rmd
Testing smaller and higher res figs

@@
 knitr::opts_knit$set(root.dir = here::here())
+
 knitr::opts_chunk$set(
   echo = FALSE, 
-  out.width = "100%", 
+  out.width = "70%", 
   warning = FALSE, 
-  message = FALSE
+  message = FALSE,
+  fig.retina = 2L
 )

### 2023-03-30 analysis/article-results.Rmd
Editing readme

@@
   message = FALSE,
-  fig.retina = 2L
+  fig.retina = 3L
 )
@@
 ) +
-  theme_minimal()
+  theme_minimal() + 
+  labs(x = "Predicted probability", y = "Observed proportion")
 ```
 
-## External validation
+## Temporal validation
 
@@
 calplot_ext
-
-ggsave(
-  calplot_ext,
-  filename = "analysis/figures/calibration-plot-ext.png",
-  width = 8,
-  height = 5,
-  units = "in",
-  dpi = 300, 
-  bg = "white"
-)
 ```
@@
   calplot_combined,
-  filename = "analysis/figures/calibration-plot-combined.png",
+  filename = "analysis/manuscript-figures/calibration-plot-combined.png",
   width = 8,
@@
 png(
-  filename = "analysis/figures/nomogram-combined.png",
+  filename = "analysis/manuscript-figures/nomogram-combined.png",
   width = 8,

### 2023-12-27 analysis/article-results.Rmd
Adding link to published url

@@
 
+## Model update: logistic recalibration
+
+```{r log-recal}
+# https://darrendahly.github.io/post/homr/
+mod_unpen <- glm(
+  model_formula,
+  family = binomial, 
+  data = imps_develop
+)
+
+df_calplot_valid$lp_unpen <- predict(mod_unpen, newdata = df_calplot_valid,
+                                     type = "link")
+
+mod_recal <- glm(
+  #Mc_FailedFemoralApproach ~ lp_unpen,
+  #Mc_FailedFemoralApproach ~ 0 + offset(lp_valid), # same model
+  #Mc_FailedFemoralApproach ~ 1 + offset(lp_valid), # Recal in the large
+  #Mc_FailedFemoralApproach ~ lp_valid, # Logistic recal
+  family = binomial, 
+  data = df_calplot_valid
+)
+
+lp_recal <- predict(mod_recal,type = "link")
+df_calplot_valid$lp_recal <- lp_recal
+probs_recal <- plogis(lp_recal)
+probs_grid <- seq(min(probs_recal), max(probs_recal), by = 0.005)
+lp_grid <- qlogis(probs_grid)
+
+calib_mod <- glm(
+  Mc_FailedFemoralApproach ~ splines::ns(lp_recal, 5), 
+  family = binomial, 
+  data = df_calplot_valid, 
+  weights = df_calplot_valid$wts
+)
+
+stacked_df <- data.frame(
+    "predicted" = probs_grid,
+    "observed" = predict(calib_mod, newdata = data.frame("lp_recal" = lp_grid), type = "response")
+)
+
+ggplot() +
+  geom_line(
+      data = stacked_df,
+      aes(x = predicted, y = observed),
+      size = 2
+    ) +
+  geom_abline(
+      intercept = 0,
+      slope = 1, col = "red", size = 1, linetype = "dashed"
+    ) +
+  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1))
+```
+
 ## Model update: combined cohort re-estimation
